---
title: "CoGrowth"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
    toc_depth: 6
    code_folding: hide
    number_sections: false
    theme: lumen

knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---
# workplace setup
```{r setup, message=FALSE, error=FALSE, warning=FALSE}
library(Hmisc)
library(vegan)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(phyloseq)
library(ape)
library(ggpubr)
library(microbiome)
library(ape)
library(rstatix)
library(ggpubr)
library(MicEco)
```


# Physiology
## Growth Rates
### Bacteria 
```{r bact-1, message=FALSE, error=FALSE, warning=FALSE}
read<-read.csv(file="datafiles/Bact_Growth_rates.csv",
               header=T,row.names=1)

boxplot(read$Growth~read$Geno)
a=aov(read$Growth~read$Geno)
summary(a)
tuk=TukeyHSD(a)
plot(tuk,las=2)

boxplot(read$Growth~read$Pop)
a=aov(read$Growth~read$Pop)
summary(a)
tuk=TukeyHSD(a)
plot(tuk,las=2)


```

### Diatoms 


### used 11 February 

```{r fcm-growth-2, dpi=300, width=5,height=6,message=FALSE, error=FALSE, warning=FALSE}
read<-read.csv(file="datafiles/TR_Growth_rates_Feb16.csv",
               header=T,row.names=1)

symnum.args <- list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf), symbols = c("****", "***", "**", "*", " "))

subset=subset(read, treatment!="RD4_YE5" & treatment!="RA5_RD4" & treatment!="RD4_NB7")

 
ggplot1=ggplot(subset, aes(x=strain, y=growth_rate, 
                   fill=culture, 
                   color=strain)) +
  geom_boxplot(alpha=.4, linewidth=0.7) +
  theme_classic(base_size = 12) + 
  scale_fill_manual(values=c('white','gray'))+

    scale_color_manual(values=c('#4D9DE0','#E15554', '#3BB273',"#7768AE")) +
  theme(plot.margin=unit(c(0,0,0,0.25),"cm"))+
# theme(legend.position="bottom",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) + 
   theme(legend.position="none",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) + 

  ylab(expression("Growth rate " * mu *  ~day^-1)) +
  guides(color = guide_legend(nrow = 2),fill = guide_legend(nrow = 2))+
  scale_x_discrete(name='') +coord_flip() +
  stat_compare_means(aes(x=strain, y=growth_rate,
                 color=factor(culture)),paired=T,  label = "p.signif", method='anova', hide.ns=TRUE,size=12)
ggplot1


```

#### with fv.fm
```{r fv-growth-a, dpi=300, width=5,height=6,message=FALSE, error=FALSE, warning=FALSE}


read<-read.csv(file="datafiles/TR_Growth_rates_Feb16.csv",
               header=T,row.names=1)
dim(read)
head(read)

read2=subset(read, strain!="NB4" & strain !="NB6" & strain !="RA4" & strain !="YC5")

gplot2=ggplot(read, 
             aes(x=strain, y=fv,
                 color=factor(strain), fill=factor(culture)))+
  geom_boxplot(alpha=.4, linewidth=0.7) +
  theme_classic() + 
  scale_fill_manual(values=c('white','gray'))+
 # scale_color_manual(values=c('#003f5c','#7a5195', '#ef5675',"#ffa600")) +
      scale_color_manual(values=c('#4D9DE0','#E15554', '#3BB273',"#7768AE")) +

    theme(legend.position="none",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) + 

 # scale_y_continuous(name='Growth Rate')+
  ylab("(Fm/Fo)/Fm") +
  guides(color = guide_legend(nrow = 2),fill = guide_legend(nrow = 2))+
  scale_x_discrete(name='')+ coord_flip()
cowplot::plot_grid(ggplot1, gplot2, 
                   rel_heights = c(1,1),
                   nrow=2,
                   align='h',
                   axis='l',
                   byrow=TRUE, labels='auto')
       #           rel_widths = c(1.7,1))

```


#### fv/fm with sign
```{r fv-growth-b, dpi=300, width=5,height=6,message=FALSE, error=FALSE, warning=FALSE}


read<-read.csv(file="datafiles/TR_Growth_rates_Feb16.csv",
               header=T,row.names=1)
dim(read)
head(read)

read2=subset(read, strain!="NB4" & strain !="NB6" & strain !="RA4" & strain !="YC5")

gplot2=ggplot(read, 
             aes(x=strain, y=fv,
                 color=factor(strain), fill=factor(culture)))+
  geom_boxplot(alpha=.4, linewidth=0.7) +
  theme_classic() + 
  scale_fill_manual(values=c('white','gray'))+
 # scale_color_manual(values=c('#003f5c','#7a5195', '#ef5675',"#ffa600")) +
      scale_color_manual(values=c('#4D9DE0','#E15554', '#3BB273',"#7768AE")) +

    theme(legend.position="none",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) + 

 # scale_y_continuous(name='Growth Rate')+
  ylab("(Fm/Fo)/Fm") +
  guides(color = guide_legend(nrow = 2),fill = guide_legend(nrow = 2))+
  scale_x_discrete(name='')+ coord_flip() +
  stat_compare_means(aes(x=strain, y=fv,
                 color=factor(culture)),paired=T,  label = "p.signif", method='anova', hide.ns=TRUE)
    

gplot2
    

cowplot::plot_grid(ggplot1, gplot2, 
                   rel_heights = c(1,1),
                   nrow=2,
                   align='h',
                   axis='l',
                   byrow=TRUE, labels='auto')
       #           rel_widths = c(1.7,1))

```



#### with terminal biomass


```{r final-growth, dpi=300, width=5,height=6,message=FALSE, error=FALSE, warning=FALSE}

counts=read.csv(file='datafiles/rotula_cellcounts.csv',header=T,row.names=1)

dim(counts)
head(counts)

gplot_counts=ggplot(counts, 
             aes(x=strain, y=X5,
                 color=factor(strain), fill=factor(bacteria)))+
  geom_boxplot(alpha=.4, linewidth=0.7) +
  theme_classic() + 
  scale_fill_manual(values=c('white','gray'))+
      scale_color_manual(values=c('#4D9DE0','#E15554', '#3BB273',"#7768AE")) +

    theme(legend.position="none",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) + 

 # scale_y_continuous(name='Growth Rate')+
  ylab("Terminal Biomass (cells/mL)") +
  guides(color = guide_legend(nrow = 2),fill = guide_legend(nrow = 2))+
  scale_x_discrete(name='')+ coord_flip() +
  stat_compare_means(aes(x=strain, y=X5,
                 color=factor(bacteria)),paired=T,  label = "p.signif", method='anova', hide.ns=TRUE)

gplot_counts

cowplot::plot_grid(ggplot1,  gplot_counts,   gplot2,
                   rel_heights = c(1,1),
                   nrow=1,
                   align='h',
                   axis='l',
                   byrow=TRUE, labels='auto')
       #           rel_widths = c(1.7,1))

```

##### stats of fv/fm




## Growth Curves

### Diatoms
```{r curves-1, message=FALSE, error=FALSE, warning=FALSE}
read=read.csv(file='datafiles/rotula_cellcounts.csv',header=T,row.names=1)
dim(read)
treatment=as.factor(read$treatment[1:48])

dim(read)
growth_data=data.frame(read[1:48,3:7])

# get mean and standard deviation
mean=aggregate(growth_data,by=list(treatment),FUN=mean,na.rm=T)
mean=t(mean[,2:6])
sd=aggregate(growth_data,by=list(treatment),FUN=sd,na.rm=T)
sd=t(sd[,2:6])
# set time
time=c(0,2,3,4,5)

doubles=read[61:69,]
treat_doub=doubles$treatment
doubles_data=doubles[,3:7]

mean_db=aggregate(doubles_data,by=list(treat_doub),FUN=mean,na.rm=T)
mean_db=t(mean_db[,2:6])
sd_db=aggregate(doubles_data,by=list(treat_doub),FUN=sd,na.rm=T)
sd_db=t(sd_db[,2:6])


```


#### Plot Subset of Strains (used)
```{r curves-5, message=FALSE, error=FALSE, warning=FALSE}

read=read.csv(file='datafiles/rotula_cellcounts.csv',header=T)
# dim(read)

growth_data=data.frame(read[,6:10])

# get mean and standard deviation
mean=aggregate(growth_data,by=list(read$treatment),FUN=mean,na.rm=T)
sd=aggregate(growth_data,by=list(read$treatment),FUN=sd,na.rm=T)
# set time
time=c(0,2,3,4,5)
#     scale_color_manual(values=c('#4D9DE0','#E15554', '#3BB273',"#7768AE")) +

{par(mfrow=c(2,2),
     mar=c(5,5,1,1))
  
  mean_ra5=as.numeric(subset(mean, Group.1=="RA5")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="RA5")[,2:6])
  #RA5
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         ylim=c(100,50000),log='y',type='b',col='#7768AE', pch=21, bg='gray',errbar.col = '#7768AE',cex=1.3,
         xlab="Time (days)",
         ylab="Diatom Concentration (cells/mL)",xaxt='n',
         yaxt='n')
  mean_ra5=as.numeric(subset(mean, Group.1=="RA5_AX")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="RA5_AX")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         col='#7768AE', pch=21, bg='white',errbar.col = '#7768AE',cex=1.3,
         type='b',add=T)
  text(1,200,"A1",font=2)
  axis(2,at=c(100,1000,12500,50000))
  axis(1, at=c(0:5),label=TRUE)
  legend("bottomright", legend=c("xenic",'axenic'), pch=21, pt.bg=c('grey','white'), bty='n')

  
  ### RD4
mean_ra5=as.numeric(subset(mean, Group.1=="RD4")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="RD4")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         ylim=c(100,50000),log='y',type='b',col='#4D9DE0', pch=21, bg='gray',errbar.col = '#4D9DE0',cex=1.3,
         xlab="Time (days)",
         ylab="Diatom Concentration (cells/mL)",xaxt='n',
         yaxt='n')
  mean_ra5=as.numeric(subset(mean, Group.1=="RD4_AX")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="RD4_AX")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         col='#4D9DE0', pch=21, bg='white',errbar.col = '#4D9DE0',cex=1.3,
         type='b',add=T)
  text(1,200,"A2",font=2)
  axis(2,at=c(100,1000,12500,50000))
  axis(1, at=c(0:5),label=TRUE)
  legend("bottomright", legend=c("xenic",'axenic'), pch=21, pt.bg=c('grey','white'), bty='n')
  
  
  
   ### RD4
mean_ra5=as.numeric(subset(mean, Group.1=="YE5")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="YE5")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         ylim=c(100,50000),log='y',type='b',col='#E15554', pch=21, bg='gray',errbar.col = '#E15554',cex=1.3,
         xlab="Time (days)",
         ylab="Diatom Concentration (cells/mL)",xaxt='n',
         yaxt='n')
  mean_ra5=as.numeric(subset(mean, Group.1=="YE5_AX")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="YE5_AX")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         col='#E15554', pch=21, bg='white',errbar.col = '#E15554',cex=1.3,
         type='b',add=T)
  text(1,200,"B1",font=2)
  axis(2,at=c(100,1000,12500,50000))
  axis(1, at=c(0:5),label=TRUE)
  legend("bottomright", legend=c("xenic",'axenic'), pch=21, pt.bg=c('grey','white'), bty='n')

  
   ### RD4
mean_ra5=as.numeric(subset(mean, Group.1=="NB7")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="NB7")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         ylim=c(100,50000),log='y',type='b',col='#3BB273', pch=21, bg='gray',errbar.col = '#3BB273',cex=1.3,
         xlab="Time (days)",
         ylab="Diatom Concentration (cells/mL)",xaxt='n',
         yaxt='n')
  mean_ra5=as.numeric(subset(mean, Group.1=="NB7_AX")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="NB7_AX")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         col='#3BB273', pch=21, bg='white',errbar.col = '#3BB273',cex=1.3,
         type='b',add=T)
  text(1,200,"C1",font=2)
  axis(2,at=c(100,1000,12500,50000))
  axis(1, at=c(0:5),label=TRUE)
  legend("bottomright", legend=c("xenic",'axenic'), pch=21, pt.bg=c('grey','white'), bty='n')
  
  }

```


###### diatom curves with ggplot
```{r}
read=read.csv(file='datafiles/rotula_cellcounts.csv',header=T)
# dim(read)

growth_data=data.frame(read[,6:10])

# get mean and standard deviation
mean=aggregate(growth_data,by=list(read$treatment),FUN=mean,na.rm=T)
sd=aggregate(growth_data,by=list(read$treatment),FUN=sd,na.rm=T)

sd_p=pivot_longer(sd, cols = X0:X5)
colnames(sd_p)=c("Diatom", "Time", "SD")

mean_p=pivot_longer(mean, cols = X0:X5)
colnames(mean_p)=c("Diatom", "Time", "Mean")
mean_p$day=parse_number(mean_p$Time)

diatom_conc=merge(mean_p, sd_p)
diatom_conc$bacteria=c(rep("xenic",5),rep("axenic",5),
                       rep("xenic",5),rep("axenic",5),
                       rep("xenic",5),rep("axenic",5),
                       rep("xenic",5),rep("axenic",5))
diatom_conc$strain=c(rep("A1", 10), rep("A2",10), rep("B1",10), rep("C1",10))

fill_bact=c('xenic'='gray','axenic'='white')
strain_colors=c("A1"="#7768AE",
                "A2"="#4D9DE0", 
                "B1"="#E15554",
                "C1"= "#3BB273" )


diatom_plot=ggplot(diatom_conc, aes(x = day, y = Mean, fill=bacteria, color=strain)) +
  geom_line(size=1) +                  # Draws the line
  geom_point( size = 3, pch=21, linewidth=1.3) +                 # Adds data points
  geom_errorbar(aes(ymin = Mean - SD, 
                    ymax = Mean + SD), 
                width = 0.1, color = "black") +
  facet_grid(~strain) +
  scale_fill_manual(values=fill_bact) +
  scale_color_manual(values=strain_colors) +
  theme_classic2(base_size=12) +
  scale_y_log10() +
  xlab("Time (days)") +
  ylab("Diatom concentration (cells/mL)") +
 theme(legend.position='none')
diatom_plot

```



```{r}

read=read.csv(file="datafiles/bact_growth.csv",header=T)
bact1=subset(read,treat !="Bact" )


bact11=mean(bact1$X3)
bact12=mean(bact1$X4)

read=read.csv(file='datafiles/rotula_cellcounts.csv',header=T)
xen=subset(read, bacteria=="xenic")

xen

mean1=mean(xen$X3)
mean2=mean(xen$X4)

bact11/mean1
bact12/mean2

```

#### bacterial curves
```{r}
read=read.csv(file="datafiles/bact_growth2.csv",header=T)
bact1=subset(read,treat !="Bact" )



mean=aggregate(bact1,by=list(bact1$treat),FUN=mean,na.rm=T)
mean$treat=NULL
mean$strain=NULL

sd=aggregate(bact1,by=list(bact1$treat),FUN=sd,na.rm=T)
sd$treat=NULL
sd$strain=NULL

sd_p=pivot_longer(sd, cols = X0:X5)
colnames(sd_p)=c("Diatom", "Time", "SD_bact")

mean_p=pivot_longer(mean, cols = X0:X5)
colnames(mean_p)=c("Diatom", "Time", "Mean_bact")
mean_p$day=parse_number(mean_p$Time)

bact_conc=merge(mean_p, sd_p)

strain_colors=c("A1"="#7768AE",
                "A2"="#4D9DE0", 
                "B1"="#E15554",
                "C1"= "#3BB273" )


bact_plot=ggplot(bact_conc, aes(x = day, y = Mean_bact,  color=Diatom)) +
  geom_line(size=1) +                  # Draws the line
  geom_point( size = 3, pch=21, linewidth=1.3) +                 # Adds data points
  geom_errorbar(aes(ymin = Mean_bact - SD_bact, 
                    ymax = Mean_bact + SD_bact), 
                width = 0.1, color = "black") +
  facet_grid(~Diatom) +
  scale_color_manual(values=strain_colors) +
  theme_classic2(base_size=12) +
  scale_y_log10() +
  xlab("Time (days)") +
  ylab("Bacterial concentration (cells/mL)") +
  theme(legend.position = 'none')
bact_plot
```


#### diatom_bact_ratio
```{r}
diatom_conc2=subset(diatom_conc, bacteria=='xenic')

both_conc=merge(diatom_conc2, bact_conc)
both_conc$ratio=(both_conc$Mean_bact/both_conc$Mean)

strain_colors=c("A1"="#7768AE",
                "A2"="#4D9DE0", 
                "B1"="#E15554",
                "C1"= "#3BB273" )


ratio_plot=ggplot(both_conc, aes(x = day, y = ratio,  color=strain)) +
  geom_line(size=1) +                  # Draws the line
  geom_point( size = 3, pch=21, linewidth=1.3) +                 # Adds data points
  facet_grid(~strain) +
  scale_color_manual(values=strain_colors) +
  theme_classic2(base_size=12) +
  scale_y_log10() +
  xlab("Time (days)") +
  ylab("Bacteria:Diatom ratio") +
  theme(legend.position = 'none')
ratio_plot
```


```{r}

cowplot::plot_grid(diatom_plot, bact_plot, ratio_plot, align='v', nrow=3)

```

## FCM Spectra

#### with lables 

```{r}
## add diatom growht rates first 
library(ggridges)
read=read.csv(file='datafiles/fcm_rotula.csv',header=T)
data1=subset(read, Temp !="22")

data1$names=paste(data1$Diatom, data1$Temp,data1$Bact)
data1$naems2=paste(data1$strain_2)



plot3=ggplot(data1, aes(RED,y=naems2, fill=factor(Treat))) +
  geom_density_ridges(alpha=0.5) +
  theme_classic(base_size=12)+
  scale_fill_manual(values=c('#631879','white','#3B4992','white', '#EE0000','white',"#008B45",'white')) +
  theme(plot.margin=unit(c(0,0,0,0),"cm"))+
  # facet_grid(~Pop) +
  theme(legend.position="none",axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  scale_x_continuous(name='692/40 mm', limits=quantile(data1$RED, c(0.01, 0.9999))) + #+facet_wrap(~Pop, scales="free")
  scale_y_discrete(name='')
plot3

ggplot(data1, aes(RED,y=naems2, fill=factor(Treat))) + 
  geom_violin(alpha=0.7, trim=TRUE, scale='area', draw_quantiles  = c(0.25, 0.5, 0.75)) +
  theme_classic(base_size=12)+
   scale_fill_manual(values=c('#631879','white','#3B4992','white', '#EE0000','white',"#008B45",'white')) +
  theme(plot.margin=unit(c(0,0,0,0),"cm")) +
  scale_x_continuous(name='692/40 mm', limits=quantile(data1$RED, c(0.01, 0.9999))) +   coord_flip() 


ggplot(data1, aes(RED,y=naems2, fill=factor(Treat))) + 
  geom_violin(alpha=0.7, trim=TRUE, draw_quantiles  = c( 0.5)) +
  theme_classic(base_size=12)+
   scale_fill_manual(values=c('#631879','white','#3B4992','white', '#EE0000','white',"#008B45",'white')) +
  scale_x_continuous(name='692/40 mm', limits=quantile(data1$RED, c(0.01, 0.9999))) +   coord_flip() 


cowplot::plot_grid(plot3,ncol=3, labels='auto')


levels(as.factor(data1$Treat))
 
ggplot(data1, aes(GREEN,y=naems2, fill=factor(Treat))) + 
  geom_density_ridges(alpha=0.5) +
  theme_classic(base_size=12)+
   scale_fill_manual(values=c('#631879','white','#3B4992','white', '#EE0000','white',"#008B45",'white')) +
  scale_x_continuous(name='692/40 mm', limits=quantile(data1$GREEN, c(0.01, 0.9999))) 


```



#### RED with FIRE 

```{r}
## add diatom growht rates first 
library(ggridges)
read=read.csv(file='datafiles/fcm_rotula.csv',header=T)
data1=subset(read, Temp !="22")

data1$names=paste(data1$Diatom, data1$Temp,data1$Bact)
data1$naems2=paste(data1$strain_2)



plot3=ggplot(data1, aes(RED,y=naems2, fill=factor(Treat))) +
  geom_density_ridges(alpha=0.5) +
  theme_classic(base_size=12)+
  scale_fill_manual(values=c('#631879','white','#3B4992','white', '#EE0000','white',"#008B45",'white')) +
  theme(plot.margin=unit(c(0,0,0,0),"cm"))+
  # facet_grid(~Pop) +
  theme(legend.position="none",axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  scale_x_continuous(name='692/40 mm', limits=quantile(data1$RED, c(0.01, 0.9999))) + #+facet_wrap(~Pop, scales="free")
  scale_y_discrete(name='')
plot3

ggplot(data1, aes(RED,y=naems2, fill=factor(Treat))) + 
  geom_violin(alpha=0.7, trim=TRUE, scale='area', draw_quantiles  = c(0.25, 0.5, 0.75)) +
  theme_classic(base_size=12)+
   scale_fill_manual(values=c('#631879','white','#3B4992','white', '#EE0000','white',"#008B45",'white')) +
  theme(plot.margin=unit(c(0,0,0,0),"cm")) +
  scale_x_continuous(name='692/40 mm', limits=quantile(data1$RED, c(0.01, 0.9999))) +   coord_flip() 


ggplot(data1, aes(RED,y=naems2, fill=factor(Treat))) + 
  geom_violin(alpha=0.7, trim=TRUE, draw_quantiles  = c( 0.5)) +
  theme_classic(base_size=12)+
   scale_fill_manual(values=c('#631879','white','#3B4992','white', '#EE0000','white',"#008B45",'white')) +
  scale_x_continuous(name='692/40 mm', limits=quantile(data1$RED, c(0.01, 0.9999))) 


cowplot::plot_grid(plot3,ncol=3, labels='auto')


levels(as.factor(data1$Treat))
 

```



#### USE ME for figure 1

```{r fv-growth-c, dpi=300, width=5,height=6,message=FALSE, error=FALSE, warning=FALSE}

read<-read.csv(file="datafiles/TR_Growth_rates_Feb16.csv",
               header=T,row.names=1)
# values_colb=c("#631879", "#3B4992", "#EE0000", "#008B45",
symnum.args <- list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf), symbols = c("****", "***", "**", "*", " "))
 subset=read
ggplot1=ggplot(subset, aes(x=strain_26, y=growth_rate, 
                   fill=culture, 
                   color=strain_26)) +
  geom_boxplot(alpha=.7) +
  theme_classic(base_size = 14) + 
  scale_fill_manual(values=c('white','gray'))+

    scale_color_manual(values=c('#631879','#3B4992', '#EE0000',"#008B45")) +
  theme(plot.margin=unit(c(0,0,0,0.25),"cm"))+
# theme(legend.position="bottom",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) + 
   theme(legend.position="none",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1)) + theme(text = element_text(size = 14)) +

  ylab(expression("Growth rate " * mu *  ~day^-1)) +
  guides(color = guide_legend(nrow = 2),fill = guide_legend(nrow = 2))+
  scale_x_discrete(name='') +coord_flip() +
  stat_compare_means(aes(x=strain_26, y=growth_rate,
                 color=factor(culture)),paired=T,  label = "p.signif", method='anova', hide.ns=TRUE, size=10)
ggplot1



read<-read.csv(file="datafiles/TR_Growth_rates_Feb16.csv",
               header=T,row.names=1)
dim(read)
head(read)
read2=subset(read, strain!="NB4" & strain !="NB6" & strain !="RA4" & strain !="YC5")



gplot2=ggplot(read, 
             aes(x=strain_26, y=fv,
                 color=factor(strain_26), fill=factor(culture)))+
  geom_boxplot(alpha=.7) +
  theme_classic(base_size=14) + 
  scale_fill_manual(values=c('white','gray'))+
 # scale_color_manual(values=c('#003f5c','#7a5195', '#ef5675',"#ffa600")) +
    scale_color_manual(values=c('#631879','#3B4992', '#EE0000',"#008B45")) +
    theme(legend.position="none",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) + 
 # scale_y_continuous(name='Growth Rate')+
  ylab("Fv/Fm") +
  guides(color = guide_legend(nrow = 2),fill = guide_legend(nrow = 2))+
  scale_x_discrete(name='')+ coord_flip() +
  stat_compare_means(aes(x=strain_26, y=fv,
                 color=factor(culture)),paired=T,  label = "p.signif", method='anova', hide.ns=TRUE, size=10)
    

gplot2
    


gplot22=ggplot(read, 
             aes(x=strain_26, y=fv,
                 color=factor(strain_26), fill=factor(culture)))+
  geom_boxplot(alpha=.7) +
  theme_classic(base_size=14) + 
  scale_fill_manual(values=c('white','gray'))+
 # scale_color_manual(values=c('#003f5c','#7a5195', '#ef5675',"#ffa600")) +
    scale_color_manual(values=c('#631879','#3B4992', '#EE0000',"#008B45")) +
    theme(legend.position="bottom",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) + 
 # scale_y_continuous(name='Growth Rate')+
  ylab("Fv/Fm") +
  guides(color = guide_legend(nrow = 2),fill = guide_legend(nrow = 2))+
  scale_x_discrete(name='')+ coord_flip() +
  stat_compare_means(aes(x=strain_26, y=fv,
                 color=factor(culture)),paired=T,  label = "p.signif", method='anova', hide.ns=TRUE)
  
gplot22

#ggsave("culture_legend.svg", plot = gplot22, width = 8, height = 5)

gplot3=ggplot(data1, aes(RED,y=naems2, color=factor(strain_2), fill=factor(Bact))) + 
  geom_violin(alpha=0.7, trim=TRUE,draw_quantiles = c(0.5)) +
  theme_classic(base_size=14)+
  ylab("") +
  scale_fill_manual(values=c('white','gray'))+
 # scale_color_manual(values=c('#003f5c','#7a5195', '#ef5675',"#ffa600")) +
    scale_color_manual(values=c('#631879','#3B4992', '#EE0000',"#008B45")) +
  scale_x_continuous(name='692/40 mm', limits=quantile(data1$RED, c(0.01, 0.9999)))  + theme(legend.position = 'none')
gplot3


data2=data1
trimmed_data <- data2 %>%
  group_by(names) %>%
  filter(
    RED >= quantile(RED, 0.025) &
    RED <= quantile(RED, 0.975)
  ) %>%
  ungroup()



gplot3=ggplot(trimmed_data, aes(RED,y=naems2, color=factor(strain_2), fill=factor(Bact))) + 
  geom_violin(alpha=0.7, trim=TRUE, draw_quantiles = c(0.5)) +
  theme_classic(base_size=14)+
  ylab("") +
  xlab("692/40 mm") + 
  scale_fill_manual(values=c('white','gray'))+
 # scale_color_manual(values=c('#003f5c','#7a5195', '#ef5675',"#ffa600")) +
    scale_color_manual(values=c('#631879','#3B4992', '#EE0000',"#008B45")) +
   theme(legend.position = 'none') 
#   stat_compare_means(aes(x=RED, y=naems2,
 #                paired=T,  label = "p.signif", method='ks.test', hide.ns=TRUE, size=10))
gplot3



clams=cowplot::plot_grid(ggplot1, gplot2, gplot3,
                   rel_heights = c(1,1),
                   nrow=1,
                   align='h',
                   axis='l',
                   byrow=TRUE, labels='auto')

clams
#ggsave("filename.svg", plot = clams, width = 8, height = 4)

```


##### stats
```{r}
read=read.csv(file='datafiles/fcm_rotula.csv',header=T)

#NbQ-B7
nbq=subset(read, strain=="NbQ-B7")

ks.test(subset(nbq, Bact=="Axenic")$FSC,subset(nbq, Bact=="Xenic")$FSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$SSC,subset(nbq, Bact=="Xenic")$SSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$RED,subset(nbq, Bact=="Xenic")$RED, alternative = 'greater', simulate.p.value=TRUE)


#NbP-YE5
nbq=subset(read, strain=="NbP-YE5")

ks.test(subset(nbq, Bact=="Axenic")$FSC,subset(nbq, Bact=="Xenic")$FSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$SSC,subset(nbq, Bact=="Xenic")$SSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$RED,subset(nbq, Bact=="Xenic")$RED, alternative = 'greater', simulate.p.value=TRUE)

#NbO-D4
nbq=subset(read, strain=="NbO-D4")

ks.test(subset(nbq, Bact=="Axenic")$FSC,subset(nbq, Bact=="Xenic")$FSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$SSC,subset(nbq, Bact=="Xenic")$SSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$RED,subset(nbq, Bact=="Xenic")$RED, alternative = 'greater', simulate.p.value=TRUE)


#NbO-A5
nbq=subset(read, strain=="NbO-A5")

ks.test(subset(nbq, Bact=="Axenic")$FSC,subset(nbq, Bact=="Xenic")$FSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$SSC,subset(nbq, Bact=="Xenic")$SSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$RED,subset(nbq, Bact=="Xenic")$RED, alternative = 'greater', simulate.p.value=TRUE)


```              
   
   
##### stats
```{r}
read=read.csv(file='datafiles/fcm_rotula.csv',header=T)

#NbQ-B7
nbq=subset(read, strain=="NbQ-B7")

ks.test(subset(nbq, Bact=="Axenic")$FSC,subset(nbq, Bact=="Xenic")$FSC, alternative = 'two.sided', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$SSC,subset(nbq, Bact=="Xenic")$SSC, alternative = 'two.sided', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$RED,subset(nbq, Bact=="Xenic")$RED, alternative = 'two.sided', simulate.p.value=TRUE)


#NbP-YE5
nbq=subset(read, strain=="NbP-YE5")

ks.test(subset(nbq, Bact=="Axenic")$FSC,subset(nbq, Bact=="Xenic")$FSC, alternative = 'two.sided', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$SSC,subset(nbq, Bact=="Xenic")$SSC, alternative = 'two.sided', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$RED,subset(nbq, Bact=="Xenic")$RED, alternative = 'two.sided', simulate.p.value=TRUE)

#NbO-D4
nbq=subset(read, strain=="NbO-D4")

ks.test(subset(nbq, Bact=="Axenic")$FSC,subset(nbq, Bact=="Xenic")$FSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$SSC,subset(nbq, Bact=="Xenic")$SSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$RED,subset(nbq, Bact=="Xenic")$RED, alternative = 'greater', simulate.p.value=TRUE)


#NbO-A5
nbq=subset(read, strain=="NbO-A5")

ks.test(subset(nbq, Bact=="Axenic")$FSC,subset(nbq, Bact=="Xenic")$FSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$SSC,subset(nbq, Bact=="Xenic")$SSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$RED,subset(nbq, Bact=="Xenic")$RED, alternative = 'greater', simulate.p.value=TRUE)


```              
              
# 16S data
```{r,echo=F,message=F}
## Chapter 3 97% OTUs
## vst transformation 
library(DESeq2)
library(phyloseq)
library(ape)
library(vegan)
library(GenomicRanges)
library(CoDaSeq)
library(zCompositions)
library(compositions)

x<-read.csv(file='datafiles/cg_feb26.csv',header=TRUE,row.names=1)
dim(x)
gloor_tab=x
OTU = otu_table(x, taxa_are_rows=T)
taxa<-read.csv(file="/Users/oliviaahern/Documents/PhD/Chapter2/CHP_3/Full_Paper/21Apr20/OTU_Taxa.csv",header=TRUE,row.names=1)
t<-as.matrix(taxa)
tax2<-tax_table(t)
tree=read.tree('datafiles/otus_tree_21apr20.tre')
map<-import_qiime_sample_data("datafiles/cg_map_17Feb26.txt")

cg = phyloseq(OTU, map,tax2,tree)

#cg_filt=subset_samples(cg, keep_2026 =='yes')

cg
cg=subset_samples(cg, keep_2026 =='yes')

phyo1 = subset_taxa(cg, !phyla=="Cyanobacteria/Chloroplast")
colSums(otu_table(phyo1))

cg=phyo1
is.rooted(phy_tree(cg))

cg_filt=subset_samples(cg, keep_2026 =='yes')

set.seed(1234)
cg_filter=rarefy_even_depth(cg_filt, sample.size=4862, rngseed=1234)
cg_filt=cg_filter
```


# General Alpha statistics 

```{r}
cg_filta=subset_samples(cg_filt, Pop2 =="COMM_T0")
cg_filta
s=specnumber(t(otu_table(cg_filta)))
s

cg_filt2=subset_samples(cg_filt, definition !="bacterioplankton")
cg_filt3=subset_samples(cg_filt2, diatom_control !="control")
cg_filt2=cg_filt3
s=specnumber(t(otu_table(cg_filt2)))
s

boxplot((s)~sample_data(cg_filt2)$treatment_26.1, las=2, xlab="", log='y')
treat2=sample_data(cg_filt2)$definition_26

spec=cbind(s, treat2)
spec2=data.frame(spec)

mean1=with(spec2, tapply(as.numeric(s), treat2, mean))
sd1=with(spec2, tapply(as.numeric(s), treat2, sd))
mean1
sd1


cg_filt2=subset_samples(cg_filt, definition !="bacterioplankton")
cg_filt3=subset_samples(cg_filt2, diatom_control =="control")
sample_data(cg_filt3)
treat=sample_data(cg_filt3)$extract
spec=cbind(specnumber(t(otu_table(cg_filt3))), treat)
spec=data.frame(spec)
with(spec, tapply(as.numeric(V1), treat, mean))
with(spec, tapply(as.numeric(V1), treat, sd))

```
### figure 1 - no control 
```{r}
cg_filt2=subset_samples(cg_filt, definition !="bacterioplankton")
cg_filt3=subset_samples(cg_filt2, diatom_control !="control")
cg_filt2=cg_filt3
s=specnumber(t(otu_table(cg_filt2)))
s

boxplot((s)~sample_data(cg_filt2)$treatment_26.1, las=2, xlab="", log='y')
treat2=sample_data(cg_filt2)$treatment_26.1

spec=cbind(s, treat2)
spec2=data.frame(spec)

spec2
#mean1=aggregate(as.numeric(s) ~ treat3, data = spec2, FUN=mean)
mean1=with(spec2, tapply(as.numeric(s), treat2, mean))
sd1=with(spec2, tapply(as.numeric(s), treat2, sd))

treatment=names(sd1)
spec_bar=cbind(mean1, sd1)
spec_bar=data.frame(spec_bar)
spec_bar$treatment=treatment

custom_order =c( "A1 Coculture","A1 Algal-cell associated", 
                 "A2 Coculture","A2 Algal-cell associated",
                 "B1 Coculture","B1 Algal-cell associated",
                 "C1 Coculture","C1 Algal-cell associated")  
spec_bar$treatment2=spec_bar$treatment
spec_bar$treatment2 <- factor(spec_bar$treatment2, levels = custom_order)
spec_bar$treatment2=spec_bar$treatment
spec_bar$treatment2 <- factor(spec_bar$treatment2, levels = custom_order)

spec_bar$treatment_ag=spec_bar$treatment
spec_bar$treatment_ag=as.factor((spec_bar$treatment_ag))
levels(spec_bar$treatment_ag)=c("Algal-cell associated","Coculture", "Algal-cell associated","Coculture", "Algal-cell associated","Coculture", "Algal-cell associated", "Coculture")
spec_bar <- spec_bar %>%
  mutate(treatment_ag = fct_rev(factor(treatment_ag)))

spec_bar$name=spec_bar$treatment
spec_bar$name=as.factor((spec_bar$name))
levels(spec_bar$name)=c("A1", "A1", "A2", "A2", "B1", "B1", "C1", "C1")

values_colb=c( "#4D9DE0", '#E15554', '#3BB273', "#7768AE", "#4D9DE0", '#E15554', '#3BB273', "#7768AE")

values_colb=c( "#E64B35", "#4DBBD5", "#00A087", "#3C5488","#E64B35", "#4DBBD5", "#00A087", "#3C5488")


values_colb=c("#631879", "#3B4992", "#EE0000", "#008B45", "#631879", "#3B4992", "#EE0000", "#008B45")


## log which i like 
ggplot(spec_bar,aes(x=name, y=mean1)) +
  geom_bar( aes(x=name, y=mean1), stat="identity", fill=values_colb, alpha=0.7) +
    geom_errorbar( aes(x=name, ymin=mean1-sd1, ymax=mean1+sd1), width=0.4, colour="black", alpha=0.9, size=.7) +
  xlab(" ") +
  ylab("Species Richness") +
#  coord_flip() +
  facet_wrap(~treatment_ag, scales='free_x') +
  theme_classic(base_size=14) +
 scale_y_continuous(trans = 'log10') +
    theme(legend.position="bottom",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=12)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

figure1=ggplot(spec_bar,aes(x=name, y=mean1)) +
  geom_bar( aes(x=name, y=mean1), stat="identity", fill=values_colb, alpha=0.7) +
    geom_errorbar( aes(x=name, ymin=mean1-sd1, ymax=mean1+sd1), width=0.4, colour="black", alpha=0.9, size=.7) +
  xlab(" ") +
  ylab("Species Richness") +
#  coord_flip() +
  facet_wrap(~treatment_ag, scales='free_x') +
  theme_classic(base_size=14) +
 scale_y_continuous(trans = 'log10') +
    theme(legend.position="bottom",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=12)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

#ggsave('figure1.svg',figure1, dpi=300, height=4, width=5)


```

# Figure 2 barplot
```{r}
cg_filter=subset_samples(cg_filt,diatom_control=="diatom")
glom=tax_glom(cg_filter, "species")

abund=transform_sample_counts(glom, function(x) (x/sum(x)))

ps_filtered=ps_prune(abund, min.samples = 0, min.reads = 0, min.abundance = 0.015)

# 3. Melt the filtered data
df_melt <- psmelt(ps_filtered)
summary(as.factor(df_melt$species))

sub1=subset(df_melt, Sample == "WGA18")
sub1[,26:29]

colors=c(
'Aestuariibacter' ="#e59d97",
'Aquibacter' ="#ffcaa5",
'Candidatus Pelagibacter' ="#e9daa2",
'Cellulophaga' ="#afb780",
'Marinobacterium' ="#f8fff1",
'Marinomonas' ="#8fc49b",
'Neptuniibacter' ="#93efe7",
'Owenweeksia' ="#ada5d9",
'Tenacibaculum' ="#f5e4ff",
'Vibrio' ="#ffc9e2",
'Wandonia' ="#c6a2ae",
'NA' ='white')


df_melt <- df_melt %>%
  mutate(definition_26 = fct_rev(factor(definition_26)))

#phy_df <- psmelt(abund)
p <- ggplot(df_melt, aes(x = Sample, y = Abundance, fill = species)) +
  theme_classic(base_size = 12) +
  xlab(" ") +
  ylab(" Relative Abundance") +
    facet_grid(~strain_26+definition_26,scales = "free", space="free") +
  geom_bar(stat = "identity",linewidth=0.3, color='black') +
  scale_fill_manual(values=colors) +
theme(axis.text.x = element_text(angle=90, size=10), legend.position = 'left') 
p
```


# nst
```{r}

library(NST)

cg_filt2=subset_samples(cg_filt, definition !="bacterioplankton")

cg_filt2

try1=subset_samples(cg_filt2, Treatment3!="Bact_T0_filtered")
try1=subset_samples(try1, Treatment3!="Bact_T0")
try1=subset_samples(try1, treatment_26!="Bacteria Control T5")

meta=cbind( sample_data(try1)$Treatment)
meta=data.frame(meta)
row.names(meta)=sample_data(try1)$X.OTU.ID
colnames(meta)=c( "Strain") # metagroup - strain and group - BA_Micro

groupie=cbind( sample_data(try1)$treatment_26)
groupie=data.frame(groupie)
row.names(groupie)=row.names(sample_data(try1))
colnames(groupie)=c( "Treatment") # metagroup - strain and group - BA_Micro
dim(groupie)

otu_tab=t(otu_table(try1))
# taxonomic nst 
nst_1=tNST(comm = otu_tab , group = groupie, abundance.weighted=FALSE, meta.group = meta, dist.method = 'jaccard')
nst_data=nst_1$index.grp
par(mar=c(10,5,1,1))
groups_ag=c('cocultured','cocultured','cocultured','cocultured',"algal-cell associated","algal-cell associated","algal-cell associated","algal-cell associated")

groups_ag=c('co','co','co','co',"algal-cell","algal-cell","algal-cell","algal-cell")
boxplot(nst_data$NST.i.jaccard~groups_ag, xlab=" ", ylab="NST", ylim=c(0,1),yaxt='n')
axis(2, at=c(0,0.25,0.5,0.75,1))
box(which='plot')

nst_ba=nst_data$NST.i.jaccard[1:4]
nst_microbiome=nst_data$NST.i.jaccard[5:8]



mean(nst_ba)
sd(nst_ba)

mean(nst_microbiome)
sd(nst_microbiome)
```




```{r}

library(NST)

cg_filt2=subset_samples(cg_filt, definition !="bacterioplankton")

cg_filt2

try1=subset_samples(cg_filt2, Treatment3!="Bact_T0_filtered")
try1=subset_samples(try1, Treatment3!="Bact_T0")
try1=subset_samples(try1, treatment_26!="Bacteria Control T5")

strain=cbind( sample_data(try1)$diatom_strain)
strain=data.frame(strain)
row.names(strain)=sample_data(try1)$X.OTU.ID
colnames(strain)=c( "Strain") # metagroup - strain and group - BA_Micro

groupie=cbind( sample_data(try1)$definition)
groupie=data.frame(groupie)
row.names(groupie)=sample_data(try1)$X.OTU.ID
colnames(groupie)=c( "Treatment") # metagroup - strain and group - BA_Micro
dim(groupie)

otu_tab=t(otu_table(try1))
# taxonomic nst 
nst_1=tNST(comm = otu_tab , group = groupie, meta.group = strain, dist.method = 'jaccard',abundance.weighted=FALSE)
nst_data_ag=nst_1$index.grp
nst_data_ag
nst_data_group=nst_data_ag$ST.i.jaccard
names(nst_data_group)=c("co", "algal-cell")
par(mar=c(10,5,1,1))
groups_ag=c('co',"algal-cell")
barplot(nst_data_group,xlab=" ", ylab="NST", ylim=c(0,1),yaxt='n')
axis(2, at=c(0,0.25,0.5,0.75,1))
box(which='plot')

```

```{r}

library(NST)

cg_filt2=subset_samples(cg_filt, definition !="bacterioplankton")

cg_filt2

try1=subset_samples(cg_filt2, Treatment3!="Bact_T0_filtered")
try1=subset_samples(try1, Treatment3!="Bact_T0")
try1=subset_samples(try1, treatment_26!="Bacteria Control T5")

strain=cbind( sample_data(try1)$diatom_strain)
strain=data.frame(strain)
row.names(strain)=sample_data(try1)$X.OTU.ID
colnames(strain)=c( "Strain") # metagroup - strain and group - BA_Micro

groupie=cbind( sample_data(try1)$definition)
groupie=data.frame(groupie)
row.names(groupie)=sample_data(try1)$X.OTU.ID
colnames(groupie)=c( "Treatment") # metagroup - strain and group - BA_Micro
dim(groupie)

otu_tab=t(otu_table(try1))
# taxonomic nst 
nst_1=tNST(comm = otu_tab , group = groupie, meta.group = strain, dist.method = 'jaccard',abundance.weighted=FALSE)
nst_data_ag=nst_1$index.grp
nst_data_ag
nst_data_group=nst_data_ag$ST.i.jaccard
names(nst_data_group)=c("co", "algal-cell")
par(mar=c(7,5,1,1))
groups_ag=c('co',"algal-cell")
barplot(nst_data_group,xlab=" ", ylab="NST", ylim=c(0,1),yaxt='n', col=c("gray60","gray40"),las=2,cex.names=1.2)
axis(2, at=c(0,0.25,0.5,0.75,1))
#box(which='plot')

```
# Beta diversity and ordinations

## all samples pcoa
```{r beta-1}

diatom=subset_samples(cg_filt, diatom_control=='diatom')

diatom2=transform_sample_counts(diatom, function(x) log(x+1))
bray=vegdist(t(otu_table(diatom2)),'bray')

p=prcomp(bray)
summary(p)

{
  
  par(mar=c(10,
            5,1,1),xpd=FALSE)
  plot(p$x[,1],p$x[,2],pch=sample_data(diatom)$pch,
     bg=as.character(sample_data(diatom)$col_26),
     xlab = "PCoA1 54 %", 
     ylab= "PCoA2 12 %")
 # axis(2, at=c(-2,0,2))
ordiellipse(p$x,          group=as.factor(sample_data(diatom)$treatment_26.1),

           kind ='sd',conf=0.8,
           label=TRUE)
legend(-2, -5,
       legend=c("NbO-A5 BA","NbO-A5 Micro", "NbO-D4 BA", "NbO-D4 Micro", "NbP-YE5 BA", "NbP-YE5 Micro", "NbQ-B7 BA",
                "NbQ-B7 Micro"),
       bty='n',
       pt.bg=c("#4D9DE0", "#4D9DE0", "#E15554", "#E15554", "#3BB273", "#3BB273", "#7768AE", "#7768AE"),

       pch=c(21,22,21,22,21,22,21,22))

}


```

## just microbiomes
```{r beta-2a}
cg_filt2=subset_samples(cg_filt, definition_26 =='Algal-cell associated')
diatom_micro=transform_sample_counts(cg_filt2, function(x) log(x+1))
bray=vegdist(t(otu_table(diatom_micro)),'bray')

p_micro=prcomp(bray)
summary(p_micro)

mod <- betadisper(bray,group=as.factor(sample_data(diatom_micro)$strain_26))
anova(mod)
permutest(mod)
TukeyHSD(mod)
plot(mod, ellipse = TRUE, hull = FALSE, conf = 0.90) # 90% data ellipse
color_micro=c(sample_data(diatom_micro)$col_26)

{
  par(mar=c(10,
            5,1,1),xpd=FALSE)
  plot(p_micro$x[,1],p_micro$x[,2],pch=sample_data(diatom_micro)$pch,
     bg=as.character(sample_data(diatom_micro)$col_26), cex=2,
     xlab = "PCoA1 32.56 %", 
     ylab= "PCoA2 11.94 %")
 # axis(2, at=c(-2,0,2))
ordiellipse(p_micro$x,          group=as.factor(sample_data(diatom_micro)$strain_26),

           kind ='sd',conf=0.8,
           label=TRUE)
legend(-1, 0,
       legend=c("A1", "A2", "B1", "C1"),
       bty='n',
       pt.bg=c("#7768AE","#4D9DE0", "#E15554",  "#3BB273" ),

       pch=c(21,22,21,22,21,22,21,22))

}


anosim(bray, sample_data(diatom_micro)$strain_26)
adonis2(bray~sample_data(diatom_micro)$strain_26)

```


## just cos
```{r beta-2}
cg_filt2=subset_samples(cg_filt, definition_26 =='Coculture')
diatom_co=transform_sample_counts(cg_filt2, function(x) log(x+1))
bray=vegdist(t(otu_table(diatom_co)),'bray')

p_co=prcomp(bray)
summary(p_co)

mod <- betadisper(bray,group=as.factor(sample_data(diatom_co)$strain_26))
anova(mod)
permutest(mod)
TukeyHSD(mod)
plot(mod, ellipse = TRUE, hull = FALSE, conf = 0.90) # 90% data ellipse

{
  par(mar=c(10,
            5,1,1),xpd=FALSE)
  plot(p_co$x[,1],p_co$x[,2],pch=21,
     bg=sample_data(diatom_co)$col_26, cex=2,
     xlab = "PCoA1 30.91 %", 
     ylab= "PCoA2 17.47 %")
 # axis(2, at=c(-2,0,2))
ordiellipse(p_co$x,          group=as.factor(sample_data(diatom_co)$strain_26),

           kind ='sd',conf=0.8,
           label=TRUE)
legend(-1, 0,
       legend=c("A1", "A2", "B1", "C1"),
       bty='n',
       pt.bg=c("#7768AE","#4D9DE0", "#E15554",  "#3BB273" ),

       pch=c(21,22,21,22,21,22,21,22))

}



```


## both pcoas together 
```{r pcoa-3}


{
  par(mar=c(10,
            5,1,1),mfrow=c(1,2),xpd=FALSE)
  
  
  plot(p_co$x[,1],p_co$x[,2],pch=sample_data(diatom_co)$pch,
     bg=sample_data(diatom_co)$col_26, cex=2,
     xlab = "PCoA1 30.91 %", 
     ylab= "PCoA2 17.47 %", xaxt='none', yaxt='none')
  axis(1, at=c(-0.3,0,0.3))
  axis(2, at=c(-0.2,0,0.2))
 # axis(2, at=c(-2,0,2))
ordiellipse(p_co$x,          group=as.factor(sample_data(diatom_co)$strain_26),

           kind ='sd',conf=0.8,
           label=TRUE)
legend(-1, 0,
       legend=c("A1", "A2", "B1", "C1"),
       bty='n',
       pt.bg=c("#7768AE","#4D9DE0", "#E15554",  "#3BB273" ),

       pch=c(21,22,21,22,21,22,21,22))

plot(p_micro$x[,1],p_micro$x[,2],pch=sample_data(diatom_micro)$pch,
     bg=as.character(sample_data(diatom_micro)$col_26), cex=2,
     xlab = "PCoA1 32.56 %", 
     ylab= "PCoA2 11.94 %", xaxt='none', yaxt='none')
  axis(1, at=c(-0.5,0,0.5))
  axis(2, at=c(-0.5,0,0.5))
 # axis(2, at=c(-2,0,2))
ordiellipse(p_micro$x,          group=as.factor(sample_data(diatom_micro)$strain_26),

           kind ='sd',conf=0.8,
           label=TRUE)

}



  par(mar=c(20,5,1,1),mfrow=c(1,2),xpd=TRUE)
  plot(p_co$x[,1],p_co$x[,2],pch=sample_data(diatom_co)$pch,
     bg=sample_data(diatom_co)$col_26, cex=2,
     xlab = "PCoA1 30.91 %", 
     ylab= "PCoA2 17.47 %", xaxt='none', yaxt='none')
  axis(1, at=c(-0.3,0,0.3))
  axis(2, at=c(-0.2,0,0.2))
 # axis(2, at=c(-2,0,2))
legend(0, -0.7,
       legend=c("A1", "A2", "B1", "C1"),
       bty='n',
       pt.bg=c("#7768AE","#4D9DE0", "#E15554",  "#3BB273" ), pch=21,
       title='strain')
legend(0, -1.5,
       legend=c("Coculture", "Algal-cell Associated"),
       bty='n',
       pt.bg=c('white' ), pch=c(21,23),
       title='treatment')
      



```


### ggplot 2
#### microbiomes 

```{r}
library(vegan)
library(ggplot2)
library(dplyr)

# ---- 1. Compute distance matrix ----
# otu_table: samples as rows, taxa/ASVs as columns (relative abundance or counts)
cg_filt2=subset_samples(cg_filt,  definition_26 =='Algal-cell associated')
diatom_co=transform_sample_counts(cg_filt2, function(x) log(x+1))
bray=vegdist(t(otu_table(diatom_co)),'bray')


# ---- 2. Run PCoA ----
pcoa_result <- prcomp(bray)
# Extract axis scores
pca_scores <- as.data.frame(pcoa_result$x[, 1:2])
colnames(pca_scores) <- c("PC1", "PC2")
pca_scores$SampleID <- rownames(pca_scores)

# % variance explained per axis
var_explained <- round((pcoa_result$sdev^2 / sum(pcoa_result$sdev^2)) * 100, 1)


strain_colors=c("A1"="#7768AE",
                "A2"="#4D9DE0", 
                "B1"="#E15554",
                "C1"= "#3BB273" )

# ---- 3. Merge with sample metadata ----
# metadata should have a SampleID column matching otu_table rownames,
# plus the variable(s) you want to facet/color by
metadata=sample_data(cg_filt2)
metadata$SampleID=metadata$X.OTU.ID
pca_scores <- pca_scores %>%
  left_join(metadata, by = "SampleID")



calc_ellipse <- function(df, npoints = 100) {
  if (nrow(df) < 3) return(NULL)  # true mathematical floor: need ≥3 points for a non-degenerate 2D covariance matrix
  center <- colMeans(df[, c("PC1", "PC2")])
  cov_mat <- cov(df[, c("PC1", "PC2")])
  
  theta <- (0:npoints) * 2 * pi / npoints
  circle <- cbind(cos(theta), sin(theta))
  
  chol_decomp <- tryCatch(chol(cov_mat), error = function(e) NULL)
  if (is.null(chol_decomp)) return(NULL)  # singular matrix (e.g. n=2, points collinear)
  
  ellipse_pts <- t(center + t(circle %*% chol_decomp) * sqrt(qchisq(0.95, df = 2)))
  data.frame(PC1 = ellipse_pts[, 1], PC2 = ellipse_pts[, 2])
}

ellipse_df <- pca_scores %>%
  group_by( strain_26) %>%
  group_modify(~ calc_ellipse(.x)) %>%
  ungroup()



# ---- 4. Plot ----
pcoa1=ggplot(pca_scores, aes(x = PC1, y = PC2, fill = strain_26)) +
  geom_point(size = 3, alpha = 0.8, pch=23) +
  labs(
    x = paste0("PC1 (", var_explained[1], "%)"),
    y = paste0("PC2 (", var_explained[2], "%)"),
    color = "Group"
  ) +
  theme_classic2(base_size = 14) +
  geom_path(data = ellipse_df, aes(x = PC1, y = PC2),
            linewidth = 0.3) +
  scale_fill_manual(values=strain_colors) + 
  theme(legend.position = 'none')
pcoa1
```

#### coculutres 



```{r}

library(vegan)
library(ggplot2)
library(dplyr)

# ---- 1. Compute distance matrix ----
# otu_table: samples as rows, taxa/ASVs as columns (relative abundance or counts)
cg_filt2=subset_samples(cg_filt,  definition_26 =='Coculture')
diatom_co=transform_sample_counts(cg_filt2, function(x) log(x+1))
bray=vegdist(t(otu_table(diatom_co)),'bray')


# ---- 2. Run PCoA ----
pcoa_result <- prcomp(bray)
# Extract axis scores
pca_scores <- as.data.frame(pcoa_result$x[, 1:2])
colnames(pca_scores) <- c("PC1", "PC2")
pca_scores$SampleID <- rownames(pca_scores)

# % variance explained per axis
var_explained <- round((pcoa_result$sdev^2 / sum(pcoa_result$sdev^2)) * 100, 1)


strain_colors=c("A1"="#7768AE",
                "A2"="#4D9DE0", 
                "B1"="#E15554", "C1"= "#3BB273" )

# ---- 3. Merge with sample metadata ----
# metadata should have a SampleID column matching otu_table rownames,
# plus the variable(s) you want to facet/color by
metadata=sample_data(cg_filt2)
metadata$SampleID=metadata$X.OTU.ID
pca_scores <- pca_scores %>%
  left_join(metadata, by = "SampleID")

calc_ellipse <- function(df, npoints = 100) {
  if (nrow(df) < 3) return(NULL)  # true mathematical floor: need ≥3 points for a non-degenerate 2D covariance matrix
  center <- colMeans(df[, c("PC1", "PC2")])
  cov_mat <- cov(df[, c("PC1", "PC2")])
  
  theta <- (0:npoints) * 2 * pi / npoints
  circle <- cbind(cos(theta), sin(theta))
  
  chol_decomp <- tryCatch(chol(cov_mat), error = function(e) NULL)
  if (is.null(chol_decomp)) return(NULL)  # singular matrix (e.g. n=2, points collinear)
  
  ellipse_pts <- t(center + t(circle %*% chol_decomp) * sqrt(qchisq(0.95, df = 2)))
  data.frame(PC1 = ellipse_pts[, 1], PC2 = ellipse_pts[, 2])
}

ellipse_df <- pca_scores %>%
  group_by( strain_26) %>%
  group_modify(~ calc_ellipse(.x)) %>%
  ungroup()

pcoa2=ggplot(pca_scores, aes(x = PC1, y = PC2, fill = strain_26)) +
  geom_point(size = 3, alpha = 0.8, pch=21) +
  labs(
    x = paste0("PC1 (", var_explained[1], "%)"),
    y = paste0("PC2 (", var_explained[2], "%)"),
    color = "Group"
  ) +
  theme_classic2(base_size = 14) +
  geom_path(data = ellipse_df, aes(x = PC1, y = PC2),
            linewidth = 0.3) +
  scale_fill_manual(values=strain_colors) + theme(legend.position = 'none')

pcoa2

```


```{r}
cows=cowplot::plot_grid(pcoa2,pcoa1,labels='auto')
#ggsave('pcoa-ggplot.svg', cows, height=4, width=10)

```


#### legend 

```{r}

library(vegan)
library(ggplot2)
library(dplyr)

# ---- 1. Compute distance matrix ----
# otu_table: samples as rows, taxa/ASVs as columns (relative abundance or counts)
cg_filt2=subset_samples(cg_filt,  definition_26 =='Coculture' |
                           definition_26 =='Algal-cell associated')
diatom_co=transform_sample_counts(cg_filt2, function(x) log(x+1))
bray=vegdist(t(otu_table(diatom_co)),'bray')


# ---- 2. Run PCoA ----
pcoa_result <- prcomp(bray)
# Extract axis scores
pca_scores <- as.data.frame(pcoa_result$x[, 1:2])
colnames(pca_scores) <- c("PC1", "PC2")
pca_scores$SampleID <- rownames(pca_scores)

# % variance explained per axis
var_explained <- round((pcoa_result$sdev^2 / sum(pcoa_result$sdev^2)) * 100, 1)


strain_colors=c("A1"="#7768AE",
                "A2"="#4D9DE0", 
                "B1"="#E15554", "C1"= "#3BB273" )

# ---- 3. Merge with sample metadata ----
# metadata should have a SampleID column matching otu_table rownames,
# plus the variable(s) you want to facet/color by
metadata=sample_data(cg_filt2)
metadata$SampleID=metadata$X.OTU.ID
pca_scores <- pca_scores %>%
  left_join(metadata, by = "SampleID")

calc_ellipse <- function(df, npoints = 100) {
  if (nrow(df) < 3) return(NULL)  # true mathematical floor: need ≥3 points for a non-degenerate 2D covariance matrix
  center <- colMeans(df[, c("PC1", "PC2")])
  cov_mat <- cov(df[, c("PC1", "PC2")])
  
  theta <- (0:npoints) * 2 * pi / npoints
  circle <- cbind(cos(theta), sin(theta))
  
  chol_decomp <- tryCatch(chol(cov_mat), error = function(e) NULL)
  if (is.null(chol_decomp)) return(NULL)  # singular matrix (e.g. n=2, points collinear)
  
  ellipse_pts <- t(center + t(circle %*% chol_decomp) * sqrt(qchisq(0.95, df = 2)))
  data.frame(PC1 = ellipse_pts[, 1], PC2 = ellipse_pts[, 2])
}

ellipse_df <- pca_scores %>%
  group_by( strain_26) %>%
  group_modify(~ calc_ellipse(.x)) %>%
  ungroup()


pca_scores_all=pca_scores


pcoa_legs=ggplot(pca_scores_all, aes(x = PC1, y = PC2, fill = strain_26, color=strain_26)) +
  geom_point(size = 3, alpha = 0.8,aes(fill=strain_26, shape=definition_26)) +
  labs(
    x = paste0("PC1 (", var_explained[1], "%)"),
    y = paste0("PC2 (", var_explained[2], "%)"),
    color = "Group"
  ) +
  facet_wrap(~definition_26) +
  theme_classic2(base_size = 14) +
  scale_fill_manual(values=strain_colors) +
  scale_color_manual(values=strain_colors)+ theme(legend.position = 'left')
#ggsave('pcoa-legend.svg', plot=pcoa_legs)


```


# ribbon plot

```{r}
library(ggalluvial)
library(ggplot2)
df_melt

cg_filter1=subset_samples(cg_filt,innoc=="innoculum" | strain_26 =="A1" | strain_26==
                            "A2" | strain_26== "B1" | strain_26 =="C1")
physeq_merged <- merge_samples(cg_filter1, group = "definition_26")
glom=tax_glom(physeq_merged, "family")

abund=transform_sample_counts(glom, function(x) (x/sum(x)))
keep_taxa <- taxa_sums(abund) / nsamples(abund) > 0.01
ps_filtered <- prune_taxa(keep_taxa, abund)
ps_filtered=ps_prune(abund, min.samples = 0, min.reads = 0, min.abundance = 0.01)
tax_table(ps_filtered)
all_melt <- psmelt(ps_filtered)

all_melt <- all_melt %>%
  group_by(Sample, family) %>%
  summarise(Abundance = sum(Abundance), .groups = "drop")
all_melt$family[is.na(all_melt$family)] <- "Other"
summary(as.factor(all_melt$family))
length(summary(as.factor(all_melt$family)))
all_melt <- all_melt %>%
  mutate(Sample = fct_relevel(Sample,
                              "Innoculum", "Coculture", "Algal-cell associated"))

all_figure=ggplot(all_melt, aes(x = Sample, y = Abundance, alluvium = family, stratum = family, fill = family)) +
  geom_flow(alpha = 0.7,na.rm=T) +
  geom_stratum() +
  theme_minimal(base_size=12) +
  theme(legend.position = 'none') +
  xlab("A1 sample flow") +
  ylab("Relative Abundance")
all_figure


library(ggalluvial)
library(ggplot2)


#A1 
cg_filter1=subset_samples(cg_filt,innoc=="innoculum" | strain_26 =="A1")
physeq_merged <- merge_samples(cg_filter1, group = "definition_26")
glom=tax_glom(physeq_merged, "family")
abund=transform_sample_counts(glom, function(x) (x/sum(x)))
a1_melt <- psmelt(abund)

keep_classes <- c("Alteromonadales", "Cytophagales", "Flavobacteriales",
                  "Gammaproteobacteria_incertae_sedis", "Methylophilales",
                  "Oceanospirillales", "Pseudomonadales", "Rhodobacterales",
                  "Rhodospirillales", "SAR11",
                  "Vibrionales", "WPS-2_genera_incertae_sedis")

a1_summed <- a1_melt %>%
  mutate(family = fct_other(family, keep = keep_classes, other_level = "Other"))

a1_summed <- a1_summed %>%
  group_by(Sample, family) %>%
  summarise(Abundance = sum(Abundance), .groups = "drop")
library(forcats)
a1_summed <- a1_summed %>%
  mutate(Sample = fct_relevel(Sample,
                              "Innoculum", "Coculture", "Algal-cell associated"))
tol_muted <- c(
  "#88CCEE", "#44AA99", "#117733", "#332288",
  "#DDCC77", "#999933", "#CC6677", "#882255",
  "#AA4499", "#DDDDDD", "#6699CC", "#888888", 'black'
)

a1_figure=ggplot(a1_summed, aes(x = Sample, y = Abundance, alluvium = family, stratum = family, fill = family)) +
  geom_flow(alpha = 0.7,na.rm=T) +
  geom_stratum() +
  theme_minimal(base_size = 12) +
  theme(legend.position = 'none') +
  xlab("A1") +
  scale_fill_manual(values = tol_muted) +
  ylab("Relative Abundance")
a1_figure

a1_figure2=ggplot(a1_summed, aes(x = Sample, y = Abundance, alluvium = family, stratum = family, fill = family)) +
  geom_flow(alpha = 0.7,na.rm=T) +
  geom_stratum() +
  theme_minimal() +
  theme(legend.position = 'bottom') +
  xlab("A1") +
  scale_fill_manual(values = tol_muted) +
  ylab("Relative Abundance")
a1_figure2

#ggsave('flowplot_legend.svg', plot=a1_figure2)

#A2
cg_filter1=subset_samples(cg_filt,innoc=="innoculum" | strain_26 =="A2")
physeq_merged <- merge_samples(cg_filter1, group = "definition_26")
glom=tax_glom(physeq_merged, "family")
abund=transform_sample_counts(glom, function(x) (x/sum(x)))
a2_melt <- psmelt(abund)

a2_summed <- a2_melt %>%
  mutate(family = fct_other(family, keep = keep_classes, other_level = "Other"))

a2_summed <- a2_summed %>%
  group_by(Sample, family) %>%
  summarise(Abundance = sum(Abundance), .groups = "drop")

library(forcats)
a2_summed <- a2_summed %>%
  mutate(Sample = fct_relevel(Sample,
                                     "Innoculum", "Coculture", "Algal-cell associated"))

a2_figure=ggplot(a2_summed, aes(x = Sample, y = Abundance, alluvium = family, stratum = family, fill = family)) +
  geom_flow(alpha = 0.7) +
  geom_stratum() +
  theme_minimal(base_size = 12) +
  theme(legend.position = 'none') +
  xlab("A2") +
  scale_fill_manual(values = tol_muted) +
  ylab("Relative Abundance")
a2_figure

#B1
cg_filter1=subset_samples(cg_filt,innoc=="innoculum" | strain_26 =="B1")
physeq_merged <- merge_samples(cg_filter1, group = "definition_26")
glom=tax_glom(physeq_merged, "family")
abund=transform_sample_counts(glom, function(x) (x/sum(x)))
b1_melt <- psmelt(abund)
b1_summed <- b1_melt %>%
  mutate(family = fct_other(family, keep = keep_classes, other_level = "Other"))

b1_summed <- b1_summed %>%
  group_by(Sample, family) %>%
  summarise(Abundance = sum(Abundance), .groups = "drop")

b1_summed <- b1_summed %>%
  mutate(Sample = fct_relevel(Sample,
                              "Innoculum", "Coculture", "Algal-cell associated"))

b1_figure=ggplot(b1_summed, aes(x = Sample, y = Abundance, alluvium = family, stratum = family, fill = family)) +
  geom_flow(alpha = 0.7) +
  geom_stratum() +
  theme_minimal(base_size = 12) +
  theme(legend.position = 'none') +
  xlab("B1") +
  scale_fill_manual(values = tol_muted) +
  ylab("Relative Abundance")
b1_figure



# C1
cg_filter1=subset_samples(cg_filt,innoc=="innoculum" | strain_26 =="C1")
physeq_merged <- merge_samples(cg_filter1, group = "definition_26")
glom=tax_glom(physeq_merged, "family")
abund=transform_sample_counts(glom, function(x) (x/sum(x)))

c1_melt <- psmelt(abund)
c1_summed <- c1_melt %>%
  mutate(family = fct_other(family, keep = keep_classes, other_level = "Other"))
c1_summed <- c1_summed %>%
  group_by(Sample, family) %>%
  summarise(Abundance = sum(Abundance), .groups = "drop")

c1_summed <- c1_summed %>%
  mutate(Sample = fct_relevel(Sample,
                              "Innoculum", "Coculture", "Algal-cell associated"))

c1_figure=ggplot(c1_summed, aes(x = Sample, y = Abundance, alluvium = family, stratum = family, fill = family)) +
  geom_flow(alpha = 0.7) +
  geom_stratum() +
  theme_minimal(base_size = 12) +
  theme(legend.position = 'none') +
  xlab("C1") +
  scale_fill_manual(values = tol_muted) +
  ylab("Relative Abundance")
c1_figure

cowplot::plot_grid(a1_figure, a2_figure, b1_figure,c1_figure)
cowww=cowplot::plot_grid(a1_figure, a2_figure, b1_figure,c1_figure,ncol =4)

#ggsave('ribbon_plot.svg', plot=cowww, dpi=300, height=4, width=16)
#cowplot::plot_grid(a1_figure, a2_figure, b1_figure,c1_figure, 
 #                  ncol=1)


```




# ribbon plot genus

```{r}
library(ggalluvial)
library(ggplot2)
df_melt

cg_filter1=subset_samples(cg_filt,innoc=="innoculum" | strain_26 =="A1" | strain_26==
                            "A2" | strain_26== "B1" | strain_26 =="C1")
cg_filter1=subset_samples(cg_filt,strain_26 =="A1" | strain_26==
                            "A2" | strain_26== "B1" | strain_26 =="C1")

physeq_merged <- merge_samples(cg_filter1, group = "definition_26")
glom=tax_glom(physeq_merged, "species")

abund=transform_sample_counts(glom, function(x) (x/sum(x)))
keep_taxa <- taxa_sums(abund) / nsamples(abund) > 0.01
ps_filtered <- prune_taxa(keep_taxa, abund)
ps_filtered=ps_prune(abund, min.samples = 0, min.reads = 0, min.abundance = 0.01)
tax_table(ps_filtered)
all_melt <- psmelt(ps_filtered)
summary(as.factor(all_melt$species))
length(summary(as.factor(all_melt$species)))

all_melt <- all_melt %>%
  group_by(Sample, species) %>%
  summarise(Abundance = sum(Abundance), .groups = "drop")


summary(as.factor(all_melt$species))
length(summary(as.factor(all_melt$species)))
all_melt <- all_melt %>%
  mutate(Sample = fct_relevel(Sample,
                              "Coculture", "Algal-cell associated"))

tol_muted_15 <- c(
  "#88CCEE", "#44AA99", "#117733", "#332288",
  "#DDCC77", "#999933", "#CC6677", "#882255",
  "#AA4499", "#6699CC", "#E69F00", "#D55E00",
  "#F0E442", "#CC79A7", "#000000"
)
tol_muted_15 <- c(
  "#88CCEE", "#44AA99", "#117733", "#332288",
  "#DDCC77", "#999933", "#CC6677", "#882255",
  "#AA4499", "#DDDDDD", "#6699CC", "#888888",
  "#E69F00", "#F0E442", "black"
)

tol_muted_14 <- c(
  "#88CCEE", "#44AA99", "#117733", "#332288",
  "#DDCC77", "#999933", "#CC6677", "#882255",
  "#AA4499", "#DDDDDD", "#6699CC", "#888888",
  "#E69F00", "black"
)

tol_muted_15 <- c(
  "#88CCEE", "#44AA99", "#117733", "#332288",
  "#DDCC77", "#999933", "#CC6677", "#882255",
  "#AA4499", "#DDDDDD", "#6699CC",
  "#E69F00", "#F0E442", "black"
)


all_melt$species <- as.factor(all_melt$species) %>%
  fct_relevel("Other", after = Inf)

all_figure=ggplot(all_melt, aes(x = Sample, y = Abundance, alluvium = species, stratum = species, fill = species)) +
  geom_flow(alpha = 0.7,na.rm=T) +
  geom_stratum() +
  theme_minimal(base_size=12) +
  theme(legend.position = 'bottom') +
  xlab("A1 sample flow") +
    scale_fill_manual(values = tol_muted_15) +
  ylab("Relative Abundance")
all_figure


summary(as.factor(all_melt$species))
length(summary(as.factor(all_melt$species)))


library(ggalluvial)
library(ggplot2)


#A1 
cg_filter1=subset_samples(cg_filt,innoc=="innoculum" | strain_26 =="A1")
physeq_merged <- merge_samples(cg_filter1, group = "definition_26")
glom=tax_glom(physeq_merged, "species")
abund=transform_sample_counts(glom, function(x) (x/sum(x)))
a1_melt <- psmelt(abund)

keep_classes <- c("Aestuariibacter", "Aquibacter","Candidatus Pelagibacter", 
                  "Cellulophaga",
                  "Marinobacterium", "Marinomonas","Neptuniibacter", "Other",
                  "Owenweeksia", "Polaribacter","Pseudohongiella",
                  "Tenacibaculum", "Vibrio",
                  "Wandonia")
a1_summed <- a1_melt %>%
  mutate(species = fct_other(species, keep = keep_classes, other_level = "Other"))

a1_summed <- a1_summed %>%
  group_by(Sample, species) %>%
  summarise(Abundance = sum(Abundance), .groups = "drop")
library(forcats)
a1_summed <- a1_summed %>%
  mutate(Sample = fct_relevel(Sample,
                              "Innoculum", "Coculture", "Algal-cell associated"))
tol_muted <- c(
  "#88CCEE", "#44AA99", "#117733", "#332288",
  "#DDCC77", "#999933", "#CC6677", "#882255",
  "#AA4499", "#DDDDDD", "#6699CC", "#888888", 'black'
)

a1_figure=ggplot(a1_summed, aes(x = Sample, y = Abundance, alluvium = species, stratum = species, fill = species)) +
  geom_flow(alpha = 0.7,na.rm=T) +
  geom_stratum() +
  theme_minimal(base_size = 12) +
  theme(legend.position = 'none') +
  xlab("A1") +
  scale_fill_manual(values = tol_muted_15) +
  ylab("Relative Abundance")
a1_figure

a1_figure2=ggplot(a1_summed, aes(x = Sample, y = Abundance, alluvium = species, stratum = species, fill = species)) +
  geom_flow(alpha = 0.7,na.rm=T) +
  geom_stratum() +
  theme_minimal() +
  theme(legend.position = 'bottom') +
  xlab("A1") +
  scale_fill_manual(values = tol_muted_15) +
  ylab("Relative Abundance")
a1_figure2

#ggsave('flowplot_legend.svg', plot=a1_figure2)

#A2
cg_filter1=subset_samples(cg_filt,innoc=="innoculum" | strain_26 =="A2")
physeq_merged <- merge_samples(cg_filter1, group = "definition_26")
glom=tax_glom(physeq_merged, "species")
abund=transform_sample_counts(glom, function(x) (x/sum(x)))
a2_melt <- psmelt(abund)

a2_summed <- a2_melt %>%
  mutate(species = fct_other(species, keep = keep_classes, other_level = "Other"))

a2_summed <- a2_summed %>%
  group_by(Sample, species) %>%
  summarise(Abundance = sum(Abundance), .groups = "drop")

library(forcats)
a2_summed <- a2_summed %>%
  mutate(Sample = fct_relevel(Sample,
                                     "Innoculum", "Coculture", "Algal-cell associated"))

a2_figure=ggplot(a2_summed, aes(x = Sample, y = Abundance, alluvium = species, stratum = species, fill = species)) +
  geom_flow(alpha = 0.7) +
  geom_stratum() +
  theme_minimal(base_size = 12) +
  theme(legend.position = 'none') +
  xlab("A2") +
  scale_fill_manual(values = tol_muted_15) +
  ylab("Relative Abundance")
a2_figure

#B1
cg_filter1=subset_samples(cg_filt,innoc=="innoculum" | strain_26 =="B1")
physeq_merged <- merge_samples(cg_filter1, group = "definition_26")
glom=tax_glom(physeq_merged, "species")
abund=transform_sample_counts(glom, function(x) (x/sum(x)))
b1_melt <- psmelt(abund)
b1_summed <- b1_melt %>%
  mutate(species = fct_other(species, keep = keep_classes, other_level = "Other"))

b1_summed <- b1_summed %>%
  group_by(Sample, species) %>%
  summarise(Abundance = sum(Abundance), .groups = "drop")

b1_summed <- b1_summed %>%
  mutate(Sample = fct_relevel(Sample,
                              "Innoculum", "Coculture", "Algal-cell associated"))

b1_figure=ggplot(b1_summed, aes(x = Sample, y = Abundance, alluvium = species, stratum = species, fill = species)) +
  geom_flow(alpha = 0.7) +
  geom_stratum() +
  theme_minimal(base_size = 12) +
  theme(legend.position = 'none') +
  xlab("B1") +
  scale_fill_manual(values = tol_muted_15) +
  ylab("Relative Abundance")
b1_figure



# C1
cg_filter1=subset_samples(cg_filt,innoc=="innoculum" | strain_26 =="C1")
physeq_merged <- merge_samples(cg_filter1, group = "definition_26")
glom=tax_glom(physeq_merged, "species")
abund=transform_sample_counts(glom, function(x) (x/sum(x)))

c1_melt <- psmelt(abund)
c1_summed <- c1_melt %>%
  mutate(species = fct_other(species, keep = keep_classes, other_level = "Other"))
c1_summed <- c1_summed %>%
  group_by(Sample, species) %>%
  summarise(Abundance = sum(Abundance), .groups = "drop")

c1_summed <- c1_summed %>%
  mutate(Sample = fct_relevel(Sample,
                              "Innoculum", "Coculture", "Algal-cell associated"))

c1_figure=ggplot(c1_summed, aes(x = Sample, y = Abundance, alluvium = species, stratum = species, fill = species)) +
  geom_flow(alpha = 0.7) +
  geom_stratum() +
  theme_minimal(base_size = 12) +
  theme(legend.position = 'none') +
  xlab("C1") +
  scale_fill_manual(values = tol_muted_15) +
  ylab("Relative Abundance")
c1_figure

cowplot::plot_grid(a1_figure, a2_figure, b1_figure,c1_figure)
cowww=cowplot::plot_grid(a1_figure, a2_figure, b1_figure,c1_figure,ncol =4)

#ggsave('ribbon_plot_genus.svg', plot=cowww, dpi=300, height=4, width=16)
#ggsave('ribbon_plot_genus_leg.svg', plot=a1_figure2, dpi=300)

#cowplot::plot_grid(a1_figure, a2_figure, b1_figure,c1_figure, 
  #                 ncol=1)


```


# diff abundance
## blanks vs non blanks

https://r-from-a-learners-perspective.readthedocs.io/en/latest/part5/

```{r}
diatom_cell=subset_samples(cg, extract=="WGA")
diatom_cell_deseq2=phyloseq_to_deseq2(diatom_cell, design=~definition)
dds <- DESeq(diatom_cell_deseq2, test="Wald", fitType="parametric")
ddsLRT <- DESeq(diatom_cell_deseq2, test="LRT", reduced= ~ 1)


res = results(dds, cooksCutoff = FALSE)
alpha = 0.001
sigtab = res[which(res$padj < alpha), ]
# sigtab = sigtab[which(abs(sigtab$log2FoldChange) > 2), ] # sometimes you want to report the taxa that had higher log2FoldChange than 2
sigtab
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(diatom_cell)[rownames(sigtab), ], "matrix"))
head(sigtab)


otu1=sigtab$strain
diatom_cell_abund=transform_sample_counts(diatom_cell, function(x) x/sum(x))
sub1=subset_taxa(diatom_cell_abund, strain %in% otu1)
plot_bar(sub1, fill="strain") +facet_wrap(~treatment_26, scales='free')


x = tapply(sigtab$log2FoldChange, sigtab$class, function(x) max(x))
x = sort(x, TRUE)
sigtab$Phylum = factor(as.character(sigtab$class), levels=names(x))

x = tapply(sigtab$log2FoldChange, sigtab$species, function(x) max(x))
x = sort(x, TRUE)
sigtab$Genus = factor(as.character(sigtab$species), levels=names(x))


ggbarplot(sigtab, x = "strain", y = "log2FoldChange",
          fill = "Genus",           # change fill color by mpg_level
          color = "white",            # Set bar border colors to white
          palette = "jco",            # jco journal color palett. see ?ggpar
          sort.val = "asc",          # Sort the value in ascending order
          sort.by.groups = FALSE,     # Don't sort inside each group
          ylab = "log2FoldChange",
          legend.title = "Phylum",
          rotate = TRUE,
          ggtheme = theme_classic()
)







```

## strains against each other 
```{r}
diatom=subset_samples(cg, diatom_control=='diatom')
diatom_cell=subset_samples(diatom, extract=="WGA")

diatom_cell_deseq2=phyloseq_to_deseq2(diatom_cell, design=~strain_26)
dds <- DESeq(diatom_cell_deseq2, test="Wald", fitType="parametric")
ddsLRT <- DESeq(diatom_cell_deseq2, test="LRT", reduced= ~ 1)

## A1 vs A2
res = results(dds, cooksCutoff = FALSE, contrast=c("strain_26", "A1", "A2"))
alpha = 0.001
sigtab_a1a2 = res[which(res$padj < alpha), ]
# sigtab = sigtab[which(abs(sigtab$log2FoldChange) > 2), ] # sometimes you want to report the taxa that had higher log2FoldChange than 2
sigtab_a1a2

sigtab_a1a2 = cbind(as(sigtab_a1a2, "data.frame"), as(tax_table(diatom_cell)[rownames(sigtab_a1a2), ], "matrix"))
head(sigtab_a1a2)
dim(sigtab_a1a2)
otu1=sigtab_a1a2$strain




## A1 vs B1
res = results(dds, cooksCutoff = FALSE, contrast=c("strain_26", "A1", "B1"))
alpha = 0.001
sigtab_a1b1 = res[which(res$padj < alpha), ]
# sigtab = sigtab[which(abs(sigtab$log2FoldChange) > 2), ] # sometimes you want to report the taxa that had higher log2FoldChange than 2
sigtab_a1b1
sigtab_a1b1 = cbind(as(sigtab_a1b1, "data.frame"), as(tax_table(diatom_cell)[rownames(sigtab_a1b1), ], "matrix"))
head(sigtab_a1b1)
dim(sigtab_a1b1)
otu2=sigtab_a1b1$strain

## A1 vs C1
res = results(dds, cooksCutoff = FALSE, contrast=c("strain_26", "A1", "C1"))
alpha = 0.001
sigtab_a1c1 = res[which(res$padj < alpha), ]
# sigtab = sigtab[which(abs(sigtab$log2FoldChange) > 2), ] # sometimes you want to report the taxa that had higher log2FoldChange than 2
sigtab_a1c1
sigtab_a1c1 = cbind(as(sigtab_a1c1, "data.frame"), as(tax_table(diatom_cell)[rownames(sigtab_a1c1), ], "matrix"))
head(sigtab_a1c1)
dim(sigtab_a1c1)
otu3=sigtab_a1c1$strain


## A2 vs B1
res = results(dds, cooksCutoff = FALSE, contrast=c("strain_26", "A2", "B1"))
alpha = 0.001
sigtab_a2b1 = res[which(res$padj < alpha), ]
# sigtab = sigtab[which(abs(sigtab$log2FoldChange) > 2), ] # sometimes you want to report the taxa that had higher log2FoldChange than 2
sigtab_a2b1
sigtab_a2b1 = cbind(as(sigtab_a2b1, "data.frame"), as(tax_table(diatom_cell)[rownames(sigtab_a2b1), ], "matrix"))
head(sigtab_a2b1)
dim(sigtab_a2b1)
otu4=sigtab_a2b1$strain

## A2 vs C1
res = results(dds, cooksCutoff = FALSE, contrast=c("strain_26", "A2", "C1"))
alpha = 0.001
sigtab_a2c1 = res[which(res$padj < alpha), ]
# sigtab = sigtab[which(abs(sigtab$log2FoldChange) > 2), ] # sometimes you want to report the taxa that had higher log2FoldChange than 2
sigtab_a2c1
sigtab_a2c1 = cbind(as(sigtab_a2c1, "data.frame"), as(tax_table(diatom_cell)[rownames(sigtab_a2c1), ], "matrix"))
head(sigtab_a2c1)
dim(sigtab_a2c1)
otu5=sigtab_a2c1$strain

## B1 vs C1
res = results(dds, cooksCutoff = FALSE, contrast=c("strain_26", "B1", "C1"))
alpha = 0.001
sigtab_b1c1 = res[which(res$padj < alpha), ]
# sigtab = sigtab[which(abs(sigtab$log2FoldChange) > 2), ] # sometimes you want to report the taxa that had higher log2FoldChange than 2
sigtab_b1c1
sigtab_b1c1 = cbind(as(sigtab_b1c1, "data.frame"), as(tax_table(diatom_cell)[rownames(sigtab_b1c1), ], "matrix"))
head(sigtab_b1c1)
dim(sigtab_b1c1)
otu6=sigtab_b1c1$strain

```


### plotting 
```{r}

llama=c(otu1,otu2,otu3,otu4,otu5,otu6)
llama_pj=subset_taxa(cg, strain %in% llama)
clamm=data.frame(tax_table(llama_pj))
levels(as.factor(clamm$class))




colors2=c("Alphaproteobacteria" = "#88CCEE",
          "Bacteroidetes_incertae_sedis" = "#44AA99",
          "Bacteroidia" = "#44AA99",
          "Betaproteobacteria" = "#117733",
          "Cytophagia"="#332288",
          "Deltaproteobacteria" ="#DDCC77",
          "Epsilonproteobacteria" = "#999933",
          "Flavobacteriia" ="#CC6677",
          "Gammaproteobacteria" ="#882255",
          "Opitutae" ="#AA4499",
          "Sphingobacteriia" ="#DDDDDD",
          "Verrucomicrobiae" = "#6699CC",
          "WPS-2_genera_incertae_sedis", "#888888")

diatom_cell=subset_samples(diatom, extract=="WGA")
diatom_cell_abund=transform_sample_counts(diatom_cell, function(x) x/sum(x))
diatom_cell_abund_otu=subset_taxa(diatom_cell_abund, strain %in% llama)
plot_bar(diatom_cell_abund_otu, fill='species')  +facet_grid(~strain_26,scales='free') + theme(legend.position = 'bottom')

p <- ggplot(sigtab_a1a2, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "Genus", title="A1 vs A2") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )  + theme(legend.position = 'none')
p



p2 <- ggplot(sigtab_a1b1, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "Genus", title="A1 vs B1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p2

p3 <- ggplot(sigtab_a1c1, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "Genus", title="A1 vs C1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p3


p4 <- ggplot(sigtab_a2b1, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "Genus", title="A2 vs B1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p4

p5 <- ggplot(sigtab_a2c1, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "Genus", title="A2 vs C1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p5


p6 <- ggplot(sigtab_b1c1, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "Genus", title="B1 vs C1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p6

cowplot::plot_grid(p,p2,p3, p4, p5, p6, nrow=2)

```
### plotting 

```{r}

llama=c(otu1,otu2,otu3,otu4,otu5,otu6)
llama_pj=subset_taxa(cg, strain %in% llama)
clamm=data.frame(tax_table(llama_pj))
levels(as.factor(clamm$class))




colors2=c("Alphaproteobacteria" = "#88CCEE",
          "Bacteroidetes_incertae_sedis" = "#44AA99",
          "Bacteroidia" = "#44AA99",
          "Betaproteobacteria" = "#117733",
          "Cytophagia"="#332288",
          "Deltaproteobacteria" ="#DDCC77",
          "Epsilonproteobacteria" = "#999933",
          "Flavobacteriia" ="#CC6677",
          "Gammaproteobacteria" ="#882255",
          "Opitutae" ="#AA4499",
          "Sphingobacteriia" ="#DDDDDD",
          "Verrucomicrobiae" = "#6699CC",
          "WPS-2_genera_incertae_sedis", "#888888")

diatom_cell=subset_samples(diatom, extract=="WGA")
diatom_cell_abund=transform_sample_counts(diatom_cell, function(x) x/sum(x))
diatom_cell_abund_otu=subset_taxa(diatom_cell_abund, strain %in% llama)
plot_bar(diatom_cell_abund_otu, fill='species')  +facet_grid(~strain_26,scales='free') + theme(legend.position = 'bottom')


genus_order <- sigtab_a1a2 %>%
  group_by(genus) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(genus)
 
sigtab_a1a2$genus <- factor(sigtab_a1a2$genus, levels = genus_order)
sigtab_a1a2$class <- factor(sigtab_a1a2$class)  # one color per taxonomic class

p <- ggplot(sigtab_a1a2, aes(x = log2FoldChange, y = genus, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "Genus", title="A1 vs A2") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )  + theme(legend.position = 'none')
p


genus_order <- sigtab_a1b1 %>%
  group_by(genus) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(genus)
 
sigtab_a1b1$genus <- factor(sigtab_a1b1$genus, levels = genus_order)
sigtab_a1b1$class <- factor(sigtab_a1b1$class)  # one color per taxonomic class

p2 <- ggplot(sigtab_a1b1, aes(x = log2FoldChange, y = genus, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "Genus", title="A1 vs B1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p2


genus_order <- sigtab_a1c1 %>%
  group_by(genus) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(genus)
 
sigtab_a1c1$genus <- factor(sigtab_a1c1$genus, levels = genus_order)
sigtab_a1c1$class <- factor(sigtab_a1c1$class)  # one color per taxonomic class

p3 <- ggplot(sigtab_a1c1, aes(x = log2FoldChange, y = genus, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "Genus", title="A1 vs C1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p3

genus_order <- sigtab_a2b1 %>%
  group_by(genus) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(genus)
 
sigtab_a2b1$genus <- factor(sigtab_a2b1$genus, levels = genus_order)
sigtab_a2b1$class <- factor(sigtab_a2b1$class)  # one color per taxonomic class
p4 <- ggplot(sigtab_a2b1, aes(x = log2FoldChange, y = genus, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "Genus", title="A2 vs B1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p4
genus_order <- sigtab_a2c1 %>%
  group_by(genus) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(genus)
 
sigtab_a2c1$genus <- factor(sigtab_a2c1$genus, levels = genus_order)
sigtab_a2c1$class <- factor(sigtab_a2c1$class)  # one color per taxonomic class
p5 <- ggplot(sigtab_a2c1, aes(x = log2FoldChange, y = genus, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "Genus", title="A2 vs C1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p5
genus_order <- sigtab_b1c1 %>%
  group_by(genus) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(genus)
 
sigtab_b1c1$genus <- factor(sigtab_b1c1$genus, levels = genus_order)
sigtab_b1c1$class <- factor(sigtab_b1c1$class)  # one color per taxonomic class

p6 <- ggplot(sigtab_b1c1, aes(x = log2FoldChange, y = genus, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "Genus", title="B1 vs C1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p6

cowplot::plot_grid(p,p2,p3, p4, p5, p6, nrow=2)

```


### plotting godo

```{r}

llama=c(otu1,otu2,otu3,otu4,otu5,otu6)
llama_pj=subset_taxa(cg, strain %in% llama)
clamm=data.frame(tax_table(llama_pj))
levels(as.factor(clamm$class))




colors2=c("Alphaproteobacteria" = "#88CCEE",
          "Bacteroidetes_incertae_sedis" = "#44AA99",
          "Bacteroidia" = "#44AA99",
          "Betaproteobacteria" = "#117733",
          "Cytophagia"="#332288",
          "Deltaproteobacteria" ="#DDCC77",
          "Epsilonproteobacteria" = "#999933",
          "Flavobacteriia" ="#CC6677",
          "Gammaproteobacteria" ="#882255",
          "Opitutae" ="#AA4499",
          "Sphingobacteriia" ="#DDDDDD",
          "Verrucomicrobiae" = "#6699CC",
          "WPS-2_genera_incertae_sedis", "#888888")

diatom_cell=subset_samples(diatom, extract=="WGA")
diatom_cell_abund=transform_sample_counts(diatom_cell, function(x) x/sum(x))
diatom_cell_abund_otu=subset_taxa(diatom_cell_abund, strain %in% llama)
plot_bar(diatom_cell_abund_otu, fill='species')  +facet_grid(~strain_26,scales='free') + theme(legend.position = 'bottom')


genus_order <- sigtab_a1a2 %>%
  group_by(species) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(species)
 
sigtab_a1a2$species <- factor(sigtab_a1a2$species, levels = genus_order)
sigtab_a1a2$class <- factor(sigtab_a1a2$class)  # one color per taxonomic class

p <- ggplot(sigtab_a1a2, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "", title="A1 vs A2") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )  + theme(legend.position = 'none')
p


genus_order <- sigtab_a1b1 %>%
  group_by(species) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(species)
 
sigtab_a1b1$species <- factor(sigtab_a1b1$species, levels = genus_order)
sigtab_a1b1$class <- factor(sigtab_a1b1$class)  # one color per taxonomic class

p2 <- ggplot(sigtab_a1b1, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "", title="A1 vs B1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p2


genus_order <- sigtab_a1c1 %>%
  group_by(species) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(species)
 
sigtab_a1c1$species <- factor(sigtab_a1c1$species, levels = genus_order)
sigtab_a1c1$class <- factor(sigtab_a1c1$class)  # one color per taxonomic class

p3 <- ggplot(sigtab_a1c1, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "", title="A1 vs C1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p3

genus_order <- sigtab_a2b1 %>%
  group_by(species) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(species)
 
sigtab_a2b1$species <- factor(sigtab_a2b1$species, levels = genus_order)
sigtab_a2b1$class <- factor(sigtab_a2b1$class)  # one color per taxonomic class
p4 <- ggplot(sigtab_a2b1, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "", title="A2 vs B1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p4


genus_order <- sigtab_a2c1 %>%
  group_by(species) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(species)
 
sigtab_a2c1$species <- factor(sigtab_a2c1$species, levels = genus_order)
sigtab_a2c1$class <- factor(sigtab_a2c1$class)  # one color per taxonomic class
p5 <- ggplot(sigtab_a2c1, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "", title="A2 vs C1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p5

genus_order <- sigtab_b1c1 %>%
  group_by(species) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(species)
 
sigtab_b1c1$species <- factor(sigtab_b1c1$species, levels = genus_order)
sigtab_b1c1$class <- factor(sigtab_b1c1$class)  # one color per taxonomic class

p6 <- ggplot(sigtab_b1c1, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "", title="B1 vs C1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p6

cowplot::plot_grid(p,p2,p3, p4, p5, p6, nrow=2, align = "hv", 
          axis = "tblr", labels='auto')


co1=cowplot::plot_grid(p,p2,p3, p4, p5, p6, nrow=2, align = "hv", 
          axis = "tblr",labels='auto')

ggsave(plot=co1, 'diff_abund.svg', dpi=300, height=9.5, width=15)

```


### plotting godo

```{r}

llama=c(otu1,otu2,otu3,otu4,otu5,otu6)
llama_pj=subset_taxa(cg, strain %in% llama)
clamm=data.frame(tax_table(llama_pj))
levels(as.factor(clamm$class))


tol_muted_15 <- c(
  "#88CCEE", "#44AA99", "#117733", "#332288",
  "#DDCC77", "#999933", "#CC6677", "#882255",
  "#AA4499", "#DDDDDD", "#6699CC",
  "#E69F00", "#F0E442", "black"
)



keep_classes <- c("Aestuariibacter" = "#88CCEE", "Aquibacter" = "#44AA99","Candidatus Pelagibacter" = "#117733", 
                  "Cellulophaga" = "#332288",
                  "Marinobacterium" = "#DDCC77", "Marinomonas" = "#999933","Neptuniibacter" = "#CC6677", "Other" = 'black',
                  "Owenweeksia" = "#882255", "Polaribacter" = "#AA4499","Pseudohongiella" = "#DDDDDD",
                  "Tenacibaculum" = "#6699CC", "Vibrio" = "#E69F00",
                  "Wandonia" ="#F0E442", "Colwellia" = 'white')


keep_classes <- c("Aestuariibacter", "Aquibacter","Candidatus Pelagibacter", 
                  "Cellulophaga",
                  "Marinobacterium", "Marinomonas","Neptuniibacter", "Other",
                  "Owenweeksia", "Polaribacter","Pseudohongiella",
                  "Tenacibaculum", "Vibrio",
                  "Wandonia")

diatom_cell=subset_samples(diatom, extract=="WGA")
diatom_cell_abund=transform_sample_counts(diatom_cell, function(x) x/sum(x))
diatom_cell_abund_otu=subset_taxa(diatom_cell_abund, strain %in% llama)
plot_bar(diatom_cell_abund_otu, fill='species')  +facet_grid(~strain_26,scales='free') + theme(legend.position = 'bottom')


genus_order <- sigtab_a1a2 %>%
  group_by(species) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(species)
 
sigtab_a1a2$species <- factor(sigtab_a1a2$species, levels = genus_order)
sigtab_a1a2$class <- factor(sigtab_a1a2$class)  # one color per taxonomic class

p <- ggplot(sigtab_a1a2, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "", title="A1 vs A2") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )  + theme(legend.position = 'none')
p


genus_order <- sigtab_a1b1 %>%
  group_by(species) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(species)
 
sigtab_a1b1$species <- factor(sigtab_a1b1$species, levels = genus_order)
sigtab_a1b1$class <- factor(sigtab_a1b1$class)  # one color per taxonomic class

p2 <- ggplot(sigtab_a1b1, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "", title="A1 vs B1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p2


genus_order <- sigtab_a1c1 %>%
  group_by(species) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(species)
 
sigtab_a1c1$species <- factor(sigtab_a1c1$species, levels = genus_order)
sigtab_a1c1$class <- factor(sigtab_a1c1$class)  # one color per taxonomic class

p3 <- ggplot(sigtab_a1c1, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "", title="A1 vs C1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p3

genus_order <- sigtab_a2b1 %>%
  group_by(species) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(species)
 
sigtab_a2b1$species <- factor(sigtab_a2b1$species, levels = genus_order)
sigtab_a2b1$class <- factor(sigtab_a2b1$class)  # one color per taxonomic class
p4 <- ggplot(sigtab_a2b1, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "", title="A2 vs B1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p4


genus_order <- sigtab_a2c1 %>%
  group_by(species) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(species)
 
sigtab_a2c1$species <- factor(sigtab_a2c1$species, levels = genus_order)
sigtab_a2c1$class <- factor(sigtab_a2c1$class)  # one color per taxonomic class
p5 <- ggplot(sigtab_a2c1, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "", title="A2 vs C1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p5

genus_order <- sigtab_b1c1 %>%
  group_by(species) %>%
  summarize(med_lfc = median(log2FoldChange, na.rm = TRUE)) %>%
  arrange(med_lfc) %>%
  pull(species)
 
sigtab_b1c1$species <- factor(sigtab_b1c1$species, levels = genus_order)
sigtab_b1c1$class <- factor(sigtab_b1c1$class)  # one color per taxonomic class

p6 <- ggplot(sigtab_b1c1, aes(x = log2FoldChange, y = species, fill = class)) +
  geom_vline(xintercept = 0, color = "grey60", linetype = "solid") +
  geom_point(size = 3, alpha = 0.9, pch=21) +
  scale_fill_manual(values = colors2, name = "Class") +
  labs(x = "log2FoldChange", y = "", title="B1 vs C1") +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    panel.grid.minor   = element_blank(),
    axis.text.y  = element_text(size = 10),
    axis.title   = element_text(face = "bold")
  )+ theme(legend.position = 'none')
p6

cowplot::plot_grid(p,p2,p3, p4, p5, p6, nrow=2, align = "hv", 
          axis = "tblr", labels='auto')


co1=cowplot::plot_grid(p,p2,p3, p4, p5, p6, nrow=2, align = "hv", 
          axis = "tblr",labels='auto')

ggsave(plot=co1, 'diff_abund.svg', dpi=300, height=9.5, width=15)

```


### plotting  bar plot

```{r}

llama=c(otu1,otu2,otu3,otu4,otu5,otu6)
llama_pj=subset_taxa(cg, strain %in% llama)
clamm=data.frame(tax_table(llama_pj))
levels(as.factor(clamm$class))

colors2=c("Alphaproteobacteria" = "#88CCEE",
          "Bacteroidetes_incertae_sedis" = "#44AA99",
          "Bacteroidia" = "#44AA99",
          "Betaproteobacteria" = "#117733",
          "Cytophagia"="#332288",
          "Deltaproteobacteria" ="#DDCC77",
          "Epsilonproteobacteria" = "#999933",
          "Flavobacteriia" ="#CC6677",
          "Gammaproteobacteria" ="#882255",
          "Opitutae" ="#AA4499",
          "Sphingobacteriia" ="#DDDDDD",
          "Verrucomicrobiae" = "#6699CC",
          "WPS-2_genera_incertae_sedis", "#888888")

diatom_cell=subset_samples(diatom, extract=="WGA")
diatom_cell_abund=transform_sample_counts(diatom_cell, function(x) x/sum(x))
diatom_cell_abund_otu=subset_taxa(diatom_cell_abund, strain %in% llama)
plot_bar(diatom_cell_abund_otu, fill='class')  +facet_grid(~strain_26,scales='free') + theme(legend.position = 'bottom') + scale_fill_manual(values=colors2)

glum=taxa_glom(diatom_cell_abund_otu, )

```