---
title: "CoGrowth"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
    toc_depth: 6
    code_folding: hide
    number_sections: false
    theme: lumen

knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---
# workplace setup
```{r setup, message=FALSE, error=FALSE, warning=FALSE}
library(Hmisc)
library(vegan)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(phyloseq)
library(ape)
library(ggpubr)
library(microbiome)
library(ape)
library(rstatix)
library(ggpubr)
```


# Physiology
## Growth Rates
### Bacteria 
```{r bact-1, message=FALSE, error=FALSE, warning=FALSE}
read<-read.csv(file="datafiles/Bact_Growth_rates.csv",
               header=T,row.names=1)

boxplot(read$Growth~read$Geno)
a=aov(read$Growth~read$Geno)
summary(a)
tuk=TukeyHSD(a)
plot(tuk,las=2)

boxplot(read$Growth~read$Pop)
a=aov(read$Growth~read$Pop)
summary(a)
tuk=TukeyHSD(a)
plot(tuk,las=2)


```

### Diatoms 

#### edited and used as of 8 jan
```{r fcm-growth-1, dpi=300, width=5,height=6,message=FALSE, error=FALSE, warning=FALSE, eval=FALSE}
read<-read.csv(file="datafiles/TR_Growth_rates_December25.csv",
               header=T,row.names=1)

subset=subset(read, treatment!="RD4_YE5" & treatment!="RA5_RD4" & treatment!="RD4_NB7")
subset=subset(read, treatment!="NB7_22" & treatment!="NB7_AX-22" & treatment!="RD4_22" & treatment!="RD4_AX-22")



growth_order=c("RA4_AX","RA4","RA5_AX","RA5","RD4_AX","RD4",
               'YC5_AX','YC5',"YE5_AX","YE5","NB4_AX","NB4","NB6_AX","NB6","NB7_AX",
               "NB7","RA5_RD4", "RD4_YE5", "RD4_NB7")

 subset$names2=paste(subset$isolate, subset$temp)

 
 
 
ggplot1=ggplot(subset, aes(x=isolate, y=growth_rate, 
                   fill=culture, 
                   color=population)) +
  geom_boxplot(alpha=.4) +
  theme_classic() + 
  scale_fill_manual(values=c('white','gray'))+
  scale_color_manual(values=c('black','navy','firebrick', 'forestgreen')) +
# scale_fill_manual(values=c("gray70",'navy',"gray70",'navy',"gray70",'navy',"gray70",'navy','gray70','firebrick','gray70','firebrick','gray70','forestgreen','gray70','forestgreen','gray70','forestgreen','gray70','forestgreen')) +
 # scale_color_manual(values=c('black','navy','black','navy','black','navy','black','navy','black','firebrick','black','firebrick','black','forestgreen','black','forestgreen','black','forestgreen','black','forestgreen' )) + 
  theme(plot.margin=unit(c(0,0,0,0.25),"cm"))+
   facet_grid(clonal~.,scales='free',space='free') +
  theme(legend.position="bottom",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) + 
 # scale_y_continuous(name='Growth Rate')+
  ylab(expression("Growth rate " * mu *  ~day^-1)) +
  guides(color = guide_legend(nrow = 2),fill = guide_legend(nrow = 2))+
  #stat_compare_means(growth_rate~treatment, group.by='group',data=subset, 'anova')+
  scale_x_discrete(name='') + coord_flip()
ggplot1


compare_means(growth_rate~treatment, group.by='isolate2',data=subset, 'anova',paired=TRUE,p.adjust.method = 'none')
test=compare_means(growth_rate~treatment, group.by='isolate2',data=subset, 'anova',paired=TRUE,p.adjust.method = 'none')
test=compare_means(growth_rate~isolate, group.by='names',data=subset, 'anova',paired=TRUE,p.adjust.method = 'none')
test=compare_means(growth_rate~clonal, group.by='isolate2',data=subset, 'anova',paired=TRUE,p.adjust.method = 'none')
test=compare_means(growth_rate~clonal, data=subset, 'anova',paired=TRUE,p.adjust.method = 'none', group.by='isolate2')
test

ggplot1+stat_pvalue_manual(data=test,label='p.adj' )


```

### used 11 February 

```{r fcm-growth-2, dpi=300, width=5,height=6,message=FALSE, error=FALSE, warning=FALSE}
read<-read.csv(file="datafiles/TR_Growth_rates_Feb16.csv",
               header=T,row.names=1)

symnum.args <- list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf), symbols = c("****", "***", "**", "*", " "))

subset=subset(read, treatment!="RD4_YE5" & treatment!="RA5_RD4" & treatment!="RD4_NB7")

 
ggplot1=ggplot(subset, aes(x=strain, y=growth_rate, 
                   fill=culture, 
                   color=strain)) +
  geom_boxplot(alpha=.4, linewidth=0.7) +
  theme_classic() + 
  scale_fill_manual(values=c('white','gray'))+

    scale_color_manual(values=c('#4D9DE0','#E15554', '#3BB273',"#7768AE")) +
  theme(plot.margin=unit(c(0,0,0,0.25),"cm"))+
# theme(legend.position="bottom",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) + 
   theme(legend.position="none",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) + 

  ylab(expression("Growth rate " * mu *  ~day^-1)) +
  guides(color = guide_legend(nrow = 2),fill = guide_legend(nrow = 2))+
  scale_x_discrete(name='') +coord_flip() +
  stat_compare_means(aes(x=strain, y=growth_rate,
                 color=factor(culture)),paired=T,  label = "p.signif", method='anova', hide.ns=TRUE)
ggplot1


```

#### with fv.fm
```{r fv-growth-a, dpi=300, width=5,height=6,message=FALSE, error=FALSE, warning=FALSE}


read<-read.csv(file="datafiles/TR_Growth_rates_Feb16.csv",
               header=T,row.names=1)
dim(read)
head(read)

read2=subset(read, strain!="NB4" & strain !="NB6" & strain !="RA4" & strain !="YC5")

gplot2=ggplot(read, 
             aes(x=strain, y=fv,
                 color=factor(strain), fill=factor(culture)))+
  geom_boxplot(alpha=.4, linewidth=0.7) +
  theme_classic() + 
  scale_fill_manual(values=c('white','gray'))+
 # scale_color_manual(values=c('#003f5c','#7a5195', '#ef5675',"#ffa600")) +
      scale_color_manual(values=c('#4D9DE0','#E15554', '#3BB273',"#7768AE")) +

    theme(legend.position="none",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) + 

 # scale_y_continuous(name='Growth Rate')+
  ylab("(Fm/Fo)/Fm") +
  guides(color = guide_legend(nrow = 2),fill = guide_legend(nrow = 2))+
  scale_x_discrete(name='')+ coord_flip()
cowplot::plot_grid(ggplot1, gplot2, 
                   rel_heights = c(1,1),
                   nrow=2,
                   align='h',
                   axis='l',
                   byrow=TRUE, labels='auto')
       #           rel_widths = c(1.7,1))

```


#### fv/fm with sign
```{r fv-growth-b, dpi=300, width=5,height=6,message=FALSE, error=FALSE, warning=FALSE}


read<-read.csv(file="datafiles/TR_Growth_rates_Feb16.csv",
               header=T,row.names=1)
dim(read)
head(read)

read2=subset(read, strain!="NB4" & strain !="NB6" & strain !="RA4" & strain !="YC5")

gplot2=ggplot(read, 
             aes(x=strain, y=fv,
                 color=factor(strain), fill=factor(culture)))+
  geom_boxplot(alpha=.4, linewidth=0.7) +
  theme_classic() + 
  scale_fill_manual(values=c('white','gray'))+
 # scale_color_manual(values=c('#003f5c','#7a5195', '#ef5675',"#ffa600")) +
      scale_color_manual(values=c('#4D9DE0','#E15554', '#3BB273',"#7768AE")) +

    theme(legend.position="none",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) + 

 # scale_y_continuous(name='Growth Rate')+
  ylab("(Fm/Fo)/Fm") +
  guides(color = guide_legend(nrow = 2),fill = guide_legend(nrow = 2))+
  scale_x_discrete(name='')+ coord_flip() +
  stat_compare_means(aes(x=strain, y=fv,
                 color=factor(culture)),paired=T,  label = "p.signif", method='anova', hide.ns=TRUE)
    

gplot2
    

cowplot::plot_grid(ggplot1, gplot2, 
                   rel_heights = c(1,1),
                   nrow=2,
                   align='h',
                   axis='l',
                   byrow=TRUE, labels='auto')
       #           rel_widths = c(1.7,1))

```
#### with divisoins


```{r fv-growth-c, dpi=300, width=5,height=6,message=FALSE, error=FALSE, warning=FALSE}

read<-read.csv(file="datafiles/TR_Growth_rates_Feb16.csv",
               header=T,row.names=1)
dim(read)
head(read)

read2=subset(read, strain!="NB4" & strain !="NB6" & strain !="RA4" & strain !="YC5")

gplot3=ggplot(read, 
             aes(x=strain, y=divisions,
                 color=factor(strain), fill=factor(culture)))+
  geom_boxplot(alpha=.4, linewidth=0.7) +
  theme_classic() + 
  scale_fill_manual(values=c('white','gray'))+
      scale_color_manual(values=c('#4D9DE0','#E15554', '#3BB273',"#7768AE")) +

    theme(legend.position="none",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) + 

 # scale_y_continuous(name='Growth Rate')+
  ylab("Number of divisions") +
  guides(color = guide_legend(nrow = 2),fill = guide_legend(nrow = 2))+
  scale_x_discrete(name='')+ coord_flip() +
  stat_compare_means(aes(x=strain, y=divisions,
                 color=factor(culture)),paired=T,  label = "p.signif", method='anova', hide.ns=TRUE)

gplot3

cowplot::plot_grid(ggplot1,  gplot3,gplot2,
                   rel_heights = c(1,1),
                   nrow=1,
                   align='h',
                   axis='l',
                   byrow=TRUE, labels='auto')
       #           rel_widths = c(1.7,1))

```

#### with terminal biomass


```{r final-growth, dpi=300, width=5,height=6,message=FALSE, error=FALSE, warning=FALSE}

counts=read.csv(file='datafiles/rotula_cellcounts.csv',header=T,row.names=1)

dim(counts)
head(counts)

gplot_counts=ggplot(counts, 
             aes(x=strain, y=X5,
                 color=factor(strain), fill=factor(bacteria)))+
  geom_boxplot(alpha=.4, linewidth=0.7) +
  theme_classic() + 
  scale_fill_manual(values=c('white','gray'))+
      scale_color_manual(values=c('#4D9DE0','#E15554', '#3BB273',"#7768AE")) +

    theme(legend.position="none",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) + 

 # scale_y_continuous(name='Growth Rate')+
  ylab("Terminal Biomass (cells/mL)") +
  guides(color = guide_legend(nrow = 2),fill = guide_legend(nrow = 2))+
  scale_x_discrete(name='')+ coord_flip() +
  stat_compare_means(aes(x=strain, y=X5,
                 color=factor(bacteria)),paired=T,  label = "p.signif", method='anova', hide.ns=TRUE)

gplot_counts

cowplot::plot_grid(ggplot1,  gplot3,gplot_counts,   gplot2,
                   rel_heights = c(1,1),
                   nrow=1,
                   align='h',
                   axis='l',
                   byrow=TRUE, labels='auto')
       #           rel_widths = c(1.7,1))

```

##### stats of fv/fm




## Growth Curves

### Diatoms
```{r curves-1, message=FALSE, error=FALSE, warning=FALSE}
read=read.csv(file='datafiles/rotula_cellcounts.csv',header=T,row.names=1)
dim(read)
treatment=as.factor(read$treatment[1:48])

dim(read)
growth_data=data.frame(read[1:48,3:7])

# get mean and standard deviation
mean=aggregate(growth_data,by=list(treatment),FUN=mean,na.rm=T)
mean=t(mean[,2:6])
sd=aggregate(growth_data,by=list(treatment),FUN=sd,na.rm=T)
sd=t(sd[,2:6])
# set time
time=c(0,2,3,4,5)

doubles=read[61:69,]
treat_doub=doubles$treatment
doubles_data=doubles[,3:7]

mean_db=aggregate(doubles_data,by=list(treat_doub),FUN=mean,na.rm=T)
mean_db=t(mean_db[,2:6])
sd_db=aggregate(doubles_data,by=list(treat_doub),FUN=sd,na.rm=T)
sd_db=t(sd_db[,2:6])


```


#### Plot Subset of Strains (used)
```{r curves-5, message=FALSE, error=FALSE, warning=FALSE}

read=read.csv(file='datafiles/rotula_cellcounts.csv',header=T)
# dim(read)

growth_data=data.frame(read[,6:10])

# get mean and standard deviation
mean=aggregate(growth_data,by=list(read$treatment),FUN=mean,na.rm=T)
sd=aggregate(growth_data,by=list(read$treatment),FUN=sd,na.rm=T)
# set time
time=c(0,2,3,4,5)
#     scale_color_manual(values=c('#4D9DE0','#E15554', '#3BB273',"#7768AE")) +

{par(mfrow=c(2,2),
     mar=c(5,5,1,1))
  
  mean_ra5=as.numeric(subset(mean, Group.1=="RA5")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="RA5")[,2:6])
  #RA5
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         ylim=c(100,50000),log='y',type='b',col='#4D9DE0', pch=21, bg='gray',errbar.col = '#4D9DE0',cex=1.3,
         xlab="Time (days)",
         ylab="Diatom Concentration (cells/mL)",xaxt='n',
         yaxt='n')
  mean_ra5=as.numeric(subset(mean, Group.1=="RA5_AX")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="RA5_AX")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         col='#4D9DE0', pch=21, bg='white',errbar.col = '#4D9DE0',cex=1.3,
         type='b',add=T)
  text(1,200,"NbO-A5",font=2)
  axis(2,at=c(100,1000,12500,50000))
  axis(1, at=c(0:5),label=TRUE)
  legend("bottomright", legend=c("xenic",'axenic'), pch=21, pt.bg=c('grey','white'), bty='n')

  
  ### RD4
mean_ra5=as.numeric(subset(mean, Group.1=="RD4")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="RD4")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         ylim=c(100,50000),log='y',type='b',col='#E15554', pch=21, bg='gray',errbar.col = '#E15554',cex=1.3,
         xlab="Time (days)",
         ylab="Diatom Concentration (cells/mL)",xaxt='n',
         yaxt='n')
  mean_ra5=as.numeric(subset(mean, Group.1=="RD4_AX")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="RD4_AX")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         col='#E15554', pch=21, bg='white',errbar.col = '#E15554',cex=1.3,
         type='b',add=T)
  text(1,200,"NbO-D4",font=2)
  axis(2,at=c(100,1000,12500,50000))
  axis(1, at=c(0:5),label=TRUE)
  legend("bottomright", legend=c("xenic",'axenic'), pch=21, pt.bg=c('grey','white'), bty='n')
  
  
  
   ### RD4
mean_ra5=as.numeric(subset(mean, Group.1=="YE5")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="YE5")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         ylim=c(100,50000),log='y',type='b',col='#3BB273', pch=21, bg='gray',errbar.col = '#3BB273',cex=1.3,
         xlab="Time (days)",
         ylab="Diatom Concentration (cells/mL)",xaxt='n',
         yaxt='n')
  mean_ra5=as.numeric(subset(mean, Group.1=="YE5_AX")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="YE5_AX")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         col='#3BB273', pch=21, bg='white',errbar.col = '#3BB273',cex=1.3,
         type='b',add=T)
  text(1,200,"NbP-YE5",font=2)
  axis(2,at=c(100,1000,12500,50000))
  axis(1, at=c(0:5),label=TRUE)
  legend("bottomright", legend=c("xenic",'axenic'), pch=21, pt.bg=c('grey','white'), bty='n')

  
   ### RD4
mean_ra5=as.numeric(subset(mean, Group.1=="NB7")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="NB7")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         ylim=c(100,50000),log='y',type='b',col='#7768AE', pch=21, bg='gray',errbar.col = '#7768AE',cex=1.3,
         xlab="Time (days)",
         ylab="Diatom Concentration (cells/mL)",xaxt='n',
         yaxt='n')
  mean_ra5=as.numeric(subset(mean, Group.1=="NB7_AX")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="NB7_AX")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         col='#7768AE', pch=21, bg='white',errbar.col = '#7768AE',cex=1.3,
         type='b',add=T)
  text(1,200,"NbQ-B7",font=2)
  axis(2,at=c(100,1000,12500,50000))
  axis(1, at=c(0:5),label=TRUE)
  legend("bottomright", legend=c("xenic",'axenic'), pch=21, pt.bg=c('grey','white'), bty='n')
  
  }

```

### bacteria (need to edit)
 
```{r bact-curve, message=FALSE, error=FALSE, warning=FALSE}
read=read.csv(file="datafiles/bact_growth.csv",header=T)

dim(read)

growth_data=data.frame(read[,3:7])
range(growth_data, na.rm=T)
# get mean and standard deviation
mean=aggregate(growth_data,by=list(read$strain),FUN=mean,na.rm=T)
sd=aggregate(growth_data,by=list(read$strain),FUN=sd,na.rm=T)
# set time
time=c(0,2,3,4,5)
#     scale_color_manual(values=c('#4D9DE0','#E15554', '#3BB273',"#7768AE")) +

{par(mfrow=c(2,3),
     mar=c(5,5,1,1))
  
  mean_ra5=as.numeric(subset(mean, Group.1=="Bacteria")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="Bacteria")[,2:6])
  #RA5
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,ylim=c(100000,1500000),
       #  ylim=c(100,50000),
         log='y',type='b',col='black', pch=21, bg='black',errbar.col = 'black',cex=1.3,
         xlab="Time (days)",
         ylab="Diatom Concentration (cells/mL)",xaxt='n')
  text(1,200,"NbO-A5",font=2)
 # axis(2,at=c(100,1000,12500,50000))
  axis(1, at=c(0:5),label=TRUE)

  
  mean_ra5=as.numeric(subset(mean, Group.1=="NbO-A5")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="NbO-A5")[,2:6])
  #RA5
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,ylim=c(100000,1500000),
         
         log='y',type='b',col='#4D9DE0', pch=21, bg='#4D9DE0',errbar.col = '#4D9DE0',cex=1.3,
         xlab="Time (days)",
         ylab="Diatom Concentration (cells/mL)",xaxt='n')
  text(1,200,"NbO-A5",font=2)
  axis(1, at=c(0:5),label=TRUE)

  
  ### RD4
mean_ra5=as.numeric(subset(mean, Group.1=="NbO-D4")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="NbO-D4")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         ylim=c(100000,1500000),log='y',type='b',col='#E15554', pch=21, bg='#E15554',errbar.col = '#E15554',cex=1.3,
         xlab="Time (days)",
         ylab="Diatom Concentration (cells/mL)",xaxt='n')
  text(1,200,"NbO-D4",font=2)
  axis(1, at=c(0:5),label=TRUE)

  
  
   ### RD4
mean_ra5=as.numeric(subset(mean, Group.1=="NbP-YE5")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="NbP-YE5")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         ylim=c(100000,1500000),log='y',type='b',col='#3BB273', pch=21, bg='#3BB273',errbar.col = '#3BB273',cex=1.3,
         xlab="Time (days)",
         ylab="Bacteria Concentration (cells/mL)",xaxt='n',
         yaxt='n')
 
  text(1,200,"NbP-YE5",font=2)
  axis(1, at=c(0:5),label=TRUE)

  
   ### RD4
mean_ra5=as.numeric(subset(mean, Group.1=="NbQ-B7")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="NbQ-B7")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         ylim=c(100000,1500000),log='y',type='b',col='#7768AE', pch=21, bg='#7768AE',errbar.col = '#7768AE',cex=1.3,
         xlab="Time (days)",
         ylab="Bacteria Concentration (cells/mL)",xaxt='n')
  text(1,200,"NbQ-B7",font=2)
  axis(1, at=c(0:5),label=TRUE)

  }

```

##### growth log 

```{r bact-curve1, message=FALSE, error=FALSE, warning=FALSE}
read=read.csv(file="datafiles/bact_growth.csv",header=T)

dim(read)

growth_data=data.frame(read[,3:7])
range(growth_data, na.rm=T)
# get mean and standard deviation
mean=aggregate(growth_data,by=list(read$strain),FUN=mean,na.rm=T)
sd=aggregate(growth_data,by=list(read$strain),FUN=sd,na.rm=T)
# set time
time=c(0,2,3,4,5)
#     scale_color_manual(values=c('#4D9DE0','#E15554', '#3BB273',"#7768AE")) +

{par(mfrow=c(1,2),
     mar=c(5,5,1,1))
  
  mean_ra5=as.numeric(subset(mean, Group.1=="Bacteria")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="Bacteria")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,ylim=c(100000,1500000),
       #  ylim=c(100,50000),
         log='y',type='b',col='black', pch=21, bg='black',errbar.col = 'black',cex=1.3,
         xlab="Time (days)",
         ylab="Bacteria Concentration (cells/mL)",xaxt='n')
    axis(1, at=c(0:5),label=TRUE)

  legend("topleft", legend=c("Bacteria", "NbO-A5", "NbO-D4", "NbP-YE5", "NbQ-B7"), pch =21 , 
         pt.bg=c('black','#4D9DE0','#E15554', '#3BB273',"#7768AE"), bty='n',
         col=c('black','#4D9DE0','#E15554', '#3BB273',"#7768AE"))
  
  mean_ra5=as.numeric(subset(mean, Group.1=="NbO-A5")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="NbO-A5")[,2:6])
  
  ## RA5 add to this one
  mean_ra5=as.numeric(subset(mean, Group.1=="NbO-A5")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="NbO-A5")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,ylim=c(100000,1500000),
         
         log='y',type='b',col='#4D9DE0', pch=21, bg='#4D9DE0',errbar.col = '#4D9DE0',cex=1.3,
         xlab="Time (days)",
         ylab="Bacteria Concentration (cells/mL)",xaxt='n')
  axis(1, at=c(0:5),label=TRUE)

  
  ### RD4
mean_ra5=as.numeric(subset(mean, Group.1=="NbO-D4")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="NbO-D4")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         ylim=c(100000,1500000),log='y',type='b',col='#E15554', pch=21, bg='#E15554',errbar.col = '#E15554',cex=1.3,
         add=T)
  
  
  
   ### RD4
mean_ra5=as.numeric(subset(mean, Group.1=="NbP-YE5")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="NbP-YE5")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         ylim=c(100000,1500000),log='y',type='b',col='#3BB273', pch=21, bg='#3BB273',errbar.col = '#3BB273',cex=1.3,
         add=T)

  
   ### RD4
mean_ra5=as.numeric(subset(mean, Group.1=="NbQ-B7")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="NbQ-B7")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         ylim=c(100000,1500000),log='y',type='b',col='#7768AE', pch=21, bg='#7768AE',errbar.col = '#7768AE',cex=1.3,
         add=T)

  }

```

#### growth non log

```{r bact-curve2, message=FALSE, error=FALSE, warning=FALSE}
read=read.csv(file="datafiles/bact_growth.csv",header=T)

dim(read)

growth_data=data.frame(read[,3:7])
range(growth_data, na.rm=T)
# get mean and standard deviation
mean=aggregate(growth_data,by=list(read$strain),FUN=mean,na.rm=T)
sd=aggregate(growth_data,by=list(read$strain),FUN=sd,na.rm=T)
# set time
time=c(0,2,3,4,5)
#     scale_color_manual(values=c('#4D9DE0','#E15554', '#3BB273',"#7768AE")) +

{par(mfrow=c(1,2),
     mar=c(5,5,1,1))
  
  mean_ra5=as.numeric(subset(mean, Group.1=="Bacteria")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="Bacteria")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,ylim=c(100000,1500000),
       #  ylim=c(100,50000),
        type='b',col='black', pch=21, bg='black',errbar.col = 'black',cex=1.3,
         xlab="Time (days)",
         ylab="Bacteria Concentration (cells/mL)",xaxt='n', yaxt='n')
    axis(1, at=c(0:5),label=TRUE)
  axis(2, at=c(100000, 500000, 1000000, 1500000))
  legend("topleft", legend=c("Bacteria", "NbO-A5", "NbO-D4", "NbP-YE5", "NbQ-B7"), pch =21 , 
         pt.bg=c('black','#4D9DE0','#E15554', '#3BB273',"#7768AE"), bty='n',
         col=c('black','#4D9DE0','#E15554', '#3BB273',"#7768AE"))
  
  
  ## RA5 add to this one
  mean_ra5=as.numeric(subset(mean, Group.1=="NbO-A5")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="NbO-A5")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,ylim=c(100000,1500000),
         
type='b',col='#4D9DE0', pch=21, bg='#4D9DE0',errbar.col = '#4D9DE0',cex=1.3,
         xlab="Time (days)",
         ylab="Bacteria Concentration (cells/mL)",xaxt='n', yaxt='n')
  axis(1, at=c(0:5),label=TRUE)
axis(2, at=c(100000, 500000, 1000000, 1500000))
  
  ### RD4
mean_ra5=as.numeric(subset(mean, Group.1=="NbO-D4")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="NbO-D4")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         ylim=c(100000,1500000),type='b',col='#E15554', pch=21, bg='#E15554',errbar.col = '#E15554',cex=1.3,
         add=T)
  
  
  
   ### RD4
mean_ra5=as.numeric(subset(mean, Group.1=="NbP-YE5")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="NbP-YE5")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         ylim=c(100000,1500000),type='b',col='#3BB273', pch=21, bg='#3BB273',errbar.col = '#3BB273',cex=1.3,
         add=T)

  
   ### RD4
mean_ra5=as.numeric(subset(mean, Group.1=="NbQ-B7")[,2:6])
  sd_ra5=as.numeric(subset(sd, Group.1=="NbQ-B7")[,2:6])
  errbar(x=time,y=mean_ra5,
         yplus=mean_ra5-sd_ra5,
         yminus=mean_ra5+sd_ra5,
         ylim=c(100000,1500000),type='b',col='#7768AE', pch=21, bg='#7768AE',errbar.col = '#7768AE',cex=1.3,
         add=T)

  }

```

## FCM Spectra

#### geom density plots gooder

```{r}
## add diatom growht rates first 
library(ggridges)
read=read.csv(file='datafiles/fcm_rotula.csv',header=T)
data1=subset(read, Temp !="22")

data1$names=paste(data1$Diatom, data1$Temp,data1$Bact)
data1$naems2=paste(data1$Diatom)




plot1=ggplot(data1, aes(FSC,y=naems2, fill=factor(Treat))) +
  geom_density_ridges(alpha=0.5) +
  theme_classic()+
  scale_fill_manual(values=c('#4D9DE0','white','#E15554','white', '#3BB273','white',"#7768AE",'white')) +
  theme(plot.margin=unit(c(0,0,0,0),"cm"))+
  # facet_grid(~Pop) +
  theme(legend.position="none",axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size=8)) + 
  scale_x_continuous(name='FSC', limits=quantile(data1$FSC, c(0.001, 0.9999)))  + scale_y_discrete(name='')
  #scale_x_discrete(name='')
plot1

plot2=ggplot(data1, aes(SSC,y=naems2, fill=factor(Treat))) +
  geom_density_ridges(alpha=0.5) +
  theme_classic()+
  scale_fill_manual(values=c('#4D9DE0','white','#E15554','white', '#3BB273','white',"#7768AE",'white')) +
  theme(plot.margin=unit(c(0,0,0,0),"cm"))+
  # facet_grid(~Pop) +
  theme(legend.position="none",axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size=8)) + 
  scale_x_continuous(name='SSC', limits=quantile(data1$FSC, c(0.001, 0.9999)))  + scale_y_discrete(name='')
  #scale_x_discrete(name='')
plot2

plot3=ggplot(data1, aes(RED,y=naems2, fill=factor(Treat))) +
  geom_density_ridges(alpha=0.5) +
  theme_classic()+
  scale_fill_manual(values=c('#4D9DE0','white','#E15554','white', '#3BB273','white',"#7768AE",'white')) +
  theme(plot.margin=unit(c(0,0,0,0),"cm"))+
  # facet_grid(~Pop) +
  theme(legend.position="none",axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size=8)) + 
  scale_x_continuous(name='692/40 mm', limits=quantile(data1$RED, c(0.01, 0.9999))) + #+facet_wrap(~Pop, scales="free")
  scale_y_discrete(name='')
plot3
cowplot::plot_grid(plot1,plot2,plot3,ncol=3)


levels(as.factor(data1$Treat))
 

```


#### with lables 

```{r}
## add diatom growht rates first 
library(ggridges)
read=read.csv(file='datafiles/fcm_rotula.csv',header=T)
data1=subset(read, Temp !="22")

data1$names=paste(data1$Diatom, data1$Temp,data1$Bact)
data1$naems2=paste(data1$Diatom)




plot1=ggplot(data1, aes(FSC,y=naems2, fill=factor(Treat))) +
  geom_density_ridges(alpha=0.5) +
  theme_classic()+
  scale_fill_manual(values=c('#4D9DE0','white','#E15554','white', '#3BB273','white',"#7768AE",'white')) +
  theme(plot.margin=unit(c(0,0,0,0),"cm"))+
  # facet_grid(~Pop) +
  theme(legend.position="none",axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size=8)) + 
  scale_x_continuous(name='FSC', limits=quantile(data1$FSC, c(0.001, 0.9999)))  + scale_y_discrete(name='') +
  stat_compare_means(aes(x=FSC, y=naems2,
                 color=factor(Treat)),paired=T,  label = "p.signif", method='anova', hide.ns=TRUE)
  #scale_x_discrete(name='')
plot1

plot2=ggplot(data1, aes(SSC,y=naems2, fill=factor(Treat))) +
  geom_density_ridges(alpha=0.5) +
  theme_classic()+
  scale_fill_manual(values=c('#4D9DE0','white','#E15554','white', '#3BB273','white',"#7768AE",'white')) +
  theme(plot.margin=unit(c(0,0,0,0),"cm"))+
  # facet_grid(~Pop) +
  theme(legend.position="none",axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size=8)) + 
  scale_x_continuous(name='SSC', limits=quantile(data1$FSC, c(0.001, 0.9999)))  + scale_y_discrete(name='')
  #scale_x_discrete(name='')
plot2

plot3=ggplot(data1, aes(RED,y=naems2, fill=factor(Treat))) +
  geom_density_ridges(alpha=0.5) +
  theme_classic()+
  scale_fill_manual(values=c('#4D9DE0','white','#E15554','white', '#3BB273','white',"#7768AE",'white')) +
  theme(plot.margin=unit(c(0,0,0,0),"cm"))+
  # facet_grid(~Pop) +
  theme(legend.position="none",axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size=8)) + 
  scale_x_continuous(name='692/40 mm', limits=quantile(data1$RED, c(0.01, 0.9999))) + #+facet_wrap(~Pop, scales="free")
  scale_y_discrete(name='')
plot3
cowplot::plot_grid(plot1,plot2,plot3,ncol=3, labels='auto')


levels(as.factor(data1$Treat))
 

```


##### stats
```{r}
read=read.csv(file='datafiles/fcm_rotula.csv',header=T)

#NbQ-B7
nbq=subset(read, strain=="NbQ-B7")

ks.test(subset(nbq, Bact=="Axenic")$FSC,subset(nbq, Bact=="Xenic")$FSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$SSC,subset(nbq, Bact=="Xenic")$SSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$RED,subset(nbq, Bact=="Xenic")$RED, alternative = 'greater', simulate.p.value=TRUE)


#NbP-YE5
nbq=subset(read, strain=="NbP-YE5")

ks.test(subset(nbq, Bact=="Axenic")$FSC,subset(nbq, Bact=="Xenic")$FSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$SSC,subset(nbq, Bact=="Xenic")$SSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$RED,subset(nbq, Bact=="Xenic")$RED, alternative = 'greater', simulate.p.value=TRUE)

#NbO-D4
nbq=subset(read, strain=="NbO-D4")

ks.test(subset(nbq, Bact=="Axenic")$FSC,subset(nbq, Bact=="Xenic")$FSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$SSC,subset(nbq, Bact=="Xenic")$SSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$RED,subset(nbq, Bact=="Xenic")$RED, alternative = 'greater', simulate.p.value=TRUE)


#NbO-A5
nbq=subset(read, strain=="NbO-A5")

ks.test(subset(nbq, Bact=="Axenic")$FSC,subset(nbq, Bact=="Xenic")$FSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$SSC,subset(nbq, Bact=="Xenic")$SSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$RED,subset(nbq, Bact=="Xenic")$RED, alternative = 'greater', simulate.p.value=TRUE)


```              
   
   
##### stats
```{r}
read=read.csv(file='datafiles/fcm_rotula.csv',header=T)

#NbQ-B7
nbq=subset(read, strain=="NbQ-B7")

ks.test(subset(nbq, Bact=="Axenic")$FSC,subset(nbq, Bact=="Xenic")$FSC, alternative = 'two.sided', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$SSC,subset(nbq, Bact=="Xenic")$SSC, alternative = 'two.sided', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$RED,subset(nbq, Bact=="Xenic")$RED, alternative = 'two.sided', simulate.p.value=TRUE)


#NbP-YE5
nbq=subset(read, strain=="NbP-YE5")

ks.test(subset(nbq, Bact=="Axenic")$FSC,subset(nbq, Bact=="Xenic")$FSC, alternative = 'two.sided', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$SSC,subset(nbq, Bact=="Xenic")$SSC, alternative = 'two.sided', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$RED,subset(nbq, Bact=="Xenic")$RED, alternative = 'two.sided', simulate.p.value=TRUE)

#NbO-D4
nbq=subset(read, strain=="NbO-D4")

ks.test(subset(nbq, Bact=="Axenic")$FSC,subset(nbq, Bact=="Xenic")$FSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$SSC,subset(nbq, Bact=="Xenic")$SSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$RED,subset(nbq, Bact=="Xenic")$RED, alternative = 'greater', simulate.p.value=TRUE)


#NbO-A5
nbq=subset(read, strain=="NbO-A5")

ks.test(subset(nbq, Bact=="Axenic")$FSC,subset(nbq, Bact=="Xenic")$FSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$SSC,subset(nbq, Bact=="Xenic")$SSC, alternative = 'greater', simulate.p.value=TRUE)
ks.test(subset(nbq, Bact=="Axenic")$RED,subset(nbq, Bact=="Xenic")$RED, alternative = 'greater', simulate.p.value=TRUE)


```              
              
# 16S data
```{r,echo=F,message=F}
## Chapter 3 97% OTUs
## vst transformation 
library(DESeq2)
library(phyloseq)
library(ape)
library(vegan)
library(GenomicRanges)
library(CoDaSeq)
library(zCompositions)
library(compositions)

x<-read.csv(file='datafiles/cg_feb26.csv',header=TRUE,row.names=1)
dim(x)
gloor_tab=x
OTU = otu_table(x, taxa_are_rows=T)
taxa<-read.csv(file="/Users/oliviaahern/Documents/PhD/Chapter2/CHP_3/Full_Paper/21Apr20/OTU_Taxa.csv",header=TRUE,row.names=1)
t<-as.matrix(taxa)
tax2<-tax_table(t)
tree=read.tree('datafiles/otus_tree_21apr20.tre')
map<-import_qiime_sample_data("datafiles/cg_map_17Feb26.txt")

cg = phyloseq(OTU, map,tax2,tree)

#cg_filt=subset_samples(cg, keep_2026 =='yes')

cg
phyo1 = subset_taxa(cg, !phyla=="Cyanobacteria/Chloroplast")
cg=phyo1
is.rooted(phy_tree(cg))

cg_filt=subset_samples(cg, keep_2026 =='yes')

#physeq_filtered <- prune_taxa(taxa_sums(cg_filt) > 0, cg_filt)


```


# General Alpha statistics 
```{r}
s=specnumber(t(otu_table(cg_filt)))

s

# spec number for 
filter=subset_samples(cg_filt,clonal2=='monoclonal' & definition =='assemblage')
filter
s2=specnumber(t(otu_table(filter)))
mean(s2)
sd(s2)

filter=subset_samples(cg_filt,Treatment2=='Bact_T5')
filter
s2=specnumber(t(otu_table(filter)))
mean(s2)
sd(s2)


filter=subset_samples(cg_filt,clonal2=='monoclonal' & definition =='microbiome')
filter
s2=specnumber(t(otu_table(filter)))
mean(s2)
sd(s2) 
```


### spcies boxplot 
```{r}
cg_filt2=subset_samples(cg_filt, definition !="bacterioplankton")
s=specnumber(t(otu_table(cg_filt2)))

boxplot((s)~sample_data(cg_filt2)$treatment_26, las=2, xlab="", log='y')
treat2=sample_data(cg_filt2)$treatment_26

spec=cbind(s, treat2)
spec2=data.frame(spec)
spec2$treat2=as.factor(spec2$treat2)
levels(spec2$treat2)
spec2$treat3=spec2$treat2
spec2$treat3=as.factor(spec2$treat3)

ggplot(spec2, aes(y=as.numeric(s), x=treat2)) +
  geom_boxplot() +
  xlab(" ") +
  ylab("Species Richness") +
  coord_flip() +
    theme(legend.position="bottom",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) +
  scale_fill_manual(values=c('gray','forestgreen','forestgreen','navy','gray','navy','firebrick'))

### spcies boxplot 
```


### spcies barplot 2 USED
```{r}
cg_filt2=subset_samples(cg_filt, definition !="bacterioplankton")
s=specnumber(t(otu_table(cg_filt2)))

boxplot((s)~sample_data(cg_filt2)$treatment_26, las=2, xlab="", log='y')
treat2=sample_data(cg_filt2)$treatment_26

spec=cbind(s, treat2)
spec2=data.frame(spec)

spec2
#mean1=aggregate(as.numeric(s) ~ treat3, data = spec2, FUN=mean)
mean1=with(spec2, tapply(as.numeric(s), treat3, mean))
sd1=with(spec2, tapply(as.numeric(s), treat3, sd))

treatment=names(sd1)
spec_bar=cbind(mean1, sd1)
spec_bar=data.frame(spec_bar)
spec_bar$treatment=treatment

#   scale_color_manual(values=c('#003f5c','#7a5195', '#ef5675',"#ffa600")) +


custom_order =c("WSW Inoculum", "Filtered Inoculum", "Bacteria Control T5", "NbO-A5 BA", "NbO-A5 Micro", "NbO-D4 BA", "NbO-D4 Micro","NbP-YE5 BA", "NbP-YE5 Micro", "NbQ-B7 BA", "NbQ-B7 Micro")  
spec_bar$treatment2=spec_bar$treatment
spec_bar$treatment2 <- factor(spec_bar$treatment2, levels = custom_order)
spec_bar$treatment2=spec_bar$treatment
spec_bar$treatment2 <- factor(spec_bar$treatment2, levels = custom_order)
spec_bar$treatment_ag=spec_bar$treatment
spec_bar$treatment_ag=as.factor((spec_bar$treatment_ag))
levels(spec_bar$treatment_ag)=c("Control", "Control", "Bacterial Assemblage", "Microbiome", "Bacterial Assemblage", "Microbiome", "Bacterial Assemblage", "Microbiome", "Bacterial Assemblage", "Microbiome", "Control")


values_colb=c("gray", "gray","gray", "#4D9DE0", '#E15554', '#3BB273', "#7768AE", "#4D9DE0", '#E15554', '#3BB273', "#7768AE")

    #scale_color_manual(values=c('#4D9DE0','#E15554', '#3BB273',"#7768AE")) +

ggplot(spec_bar,aes(x=treatment, y=mean1)) +
  geom_bar( aes(x=treatment, y=mean1), stat="identity", fill=values_colb, alpha=0.7) +
    geom_errorbar( aes(x=treatment, ymin=mean1-sd1, ymax=mean1+sd1), width=0.4, colour="black", alpha=0.9, size=.7) +
  xlab(" ") +
  ylab("Species Richness") +
#  coord_flip() +
  facet_wrap(~treatment_ag, scales='free_x') +
  theme_classic() +
 scale_y_continuous(trans = 'log10') +
    theme(legend.position="bottom",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))



ggplot(spec_bar,aes(x=treatment, y=mean1)) +
  geom_bar( aes(x=treatment, y=mean1), stat="identity", fill=values_colb, alpha=0.7) +
    geom_errorbar( aes(x=treatment, ymin=mean1-sd1, ymax=mean1+sd1), width=0.4, colour="black", alpha=0.9, size=.7) +
  xlab(" ") +
  ylab("Species Richness") +
#  coord_flip() +
  facet_wrap(~treatment_ag, scales='free') +
  theme_classic() +
# scale_y_continuous(trans = 'log10') +
    theme(legend.position="bottom",axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size=10)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

```

### spcies barplot 2 USED
```{r}
cg_filter=subset_samples(cg_filt,definition=="microbiome")
glom=tax_glom(cg_filter, "family")

write.csv(tax_table(glom), 'glomtax.csv')
write.csv(otu_table(glom), 'glomtaxotu.csv')

abund=transform_sample_counts(glom, function(x) (x/sum(x)))
plot_bar(abund, fill='family') + theme(legend.position="none")

colors_family=c("Alteromonadales" = "#e1c2aa",
                "Flavobacteriales" = "#bed8b5", 
                "Oceanospirillales" = "#a2d5d6",
                "SAR11" = "#88aee1", 
                "Vibrionales" = "#b3c4e2", 
                "NA" = 'gray')

keep_taxa <- taxa_sums(abund) / nsamples(abund) > 0.01
ps_filtered <- prune_taxa(keep_taxa, abund)
ps_filtered=ps_prune(abund, min.samples = 0, min.reads = 0, min.abundance = 0.05)

# 3. Melt the filtered data
df_melt <- psmelt(ps_filtered)

#phy_df <- psmelt(abund)
p <- ggplot(df_melt, aes(x = Sample, y = Abundance, fill = family)) +
  theme_classic() +
  xlab(" ") +
    facet_grid(~ diatom_strain, space = "free", scales = "free") +
  geom_bar(stat = "identity") + 
  scale_fill_manual(values=colors_family) +
theme(axis.text.x = element_text(angle=90, size=12), legend.position = 'bottom')
p
```

```{r-barpotag}
cg_filter=subset_samples(cg_filt,diatom_control=="diatom")
glom=tax_glom(cg_filter, "class")

abund=transform_sample_counts(glom, function(x) (x/sum(x)))
plot_bar(abund, fill='class') + theme(legend.position="none")

ps_filtered=MicEco::ps_prune(abund, min.samples = 0, min.reads = 0, min.abundance = 0.01)

# 3. Melt the filtered data
df_melt <- psmelt(ps_filtered)

#phy_df <- psmelt(abund)
p <- ggplot(df_melt, aes(x = Sample, y = Abundance, fill = class)) +
  theme_classic() +
  xlab(" ") +
    facet_grid(~diatom_strain+extract,scales = "free", space="free") +
  geom_bar(stat = "identity") +
theme(axis.text.x = element_text(angle=90, size=12), legend.position = 'bottom') 
p

### tray with family 
glom=tax_glom(cg_filter, "family")

abund=transform_sample_counts(glom, function(x) (x/sum(x)))
plot_bar(abund, fill='class') + theme(legend.position="none")

ps_filtered=MicEco::ps_prune(abund, min.samples = 0, min.reads = 0, min.abundance = 0.05)

# 3. Melt the filtered data
df_melt <- psmelt(ps_filtered)

#phy_df <- psmelt(abund)
p <- ggplot(df_melt, aes(x = Sample, y = Abundance, fill = family)) +
  theme_classic() +
  xlab(" ") +
    facet_grid(~diatom_strain+extract,scales = "free", space="free") +
  geom_bar(stat = "identity",linewidth=0.3, color='black') +
    scale_fill_manual(values=colors_family) +

theme(axis.text.x = element_text(angle=90, size=10), legend.position = 'bottom') 
p
```
# circos ?
```{r}
library(MiscMetabar)
?circle_pq
cg_filt2=subset_samples(cg_filt, definition !="bacterioplankton")

cg_filt2

try1=subset_samples(cg_filt2, Treatment3!="Bact_T0_filtered")
try1=subset_samples(try1, Treatment3!="Bact_T0")
try1=subset_samples(try1, treatment_26!="Bacteria Control T5")


palette1=c("#c95466",
"#cb683a",
"#ad963e",
"#60a75c",
"#6394cc",
"#8663cb",
"#c177af",
"#ca47a0", 'gray')

palette1=c("Vibrionales"="#f1d266",
"Flavobacteriales"="#4064da",
"Oceanospirillales"="#ff8cd7",
"Alteromonadales"="#620044", "SAR 11" = "gray" , "NbO-A5 BA" = "#4D9DE0",
"NbO-A5 Micro" = "#4D9DE0",
"NbO-D4 BA" = "#E15554",
"NbO-D4 Micro" = "#E15554",
"NbP-YE5 BA" = "#3BB273",
"NbP-YE5 Micro" = "#3BB273",
"NbQ-B7 BA" = "#7768AE",
"NbQ-B7 Micro" = "#7768AE")
('#4D9DE0','white','#E15554','white', '#3BB273','white',"#7768AE",'white')

row_colors=c( "#4D9DE0", "#4D9DE0","#E15554", "#E15554","#3BB273", "#3BB273", "#7768AE", "#7768AE")
#merge=merge_samples(try1, "treatment_26")
abund=transform_sample_counts(try1, function(x) (x/sum(x)))
subset_family = subset_taxa(try1, family %in% c("Flavobacteriales", "Oceanospirillales", "Alteromonadales", "Vibrionales", "SAR 11") )
circle_pq(subset_family, fact="treatment_26", taxa = "family",
          min_prop_mod=0.001, min_prop_tax=0.001, log10trans=TRUE, add_nb_seq=TRUE, grid_col=palette1, row_col = row_colors)
circle_pq(subset_family, fact="treatment_26", taxa = "family",
          min_prop_mod=0.001, min_prop_tax=0.001, log10trans=TRUE, add_nb_seq=FALSE, grid_col=palette1, row_col = row_colors)


circle_pq(subset_family, fact="treatment_26", taxa = "family",
          min_prop_mod=0.001, min_prop_tax=0.001,  add_nb_seq=TRUE, grid_col=palette1, row_col = row_colors, log10trans = FALSE, rarefy=TRUE)


```


# nst
```{r}

library(NST)

cg_filt2=subset_samples(cg_filt, definition !="bacterioplankton")

cg_filt2

try1=subset_samples(cg_filt2, Treatment3!="Bact_T0_filtered")
try1=subset_samples(try1, Treatment3!="Bact_T0")
try1=subset_samples(try1, treatment_26!="Bacteria Control T5")

meta=cbind( sample_data(try1)$Treatment)
meta=data.frame(meta)
row.names(meta)=sample_data(try1)$X.OTU.ID
colnames(meta)=c( "Strain") # metagroup - strain and group - BA_Micro

groupie=cbind( sample_data(try1)$treatment_26)
groupie=data.frame(groupie)
row.names(groupie)=row.names(sample_data(try1))
colnames(groupie)=c( "Treatment") # metagroup - strain and group - BA_Micro
dim(groupie)

otu_tab=t(otu_table(try1))
# taxonomic nst 
nst_1=tNST(comm = otu_tab , group = groupie, abundance.weighted=FALSE, meta.group = meta, dist.method = 'jaccard')
nst_data=nst_1$index.grp
par(mar=c(10,5,1,1))
groups_ag=c('cocultured','cocultured','cocultured','cocultured',"algal-cell associated","algal-cell associated","algal-cell associated","algal-cell associated")

groups_ag=c('co','co','co','co',"algal-cell","algal-cell","algal-cell","algal-cell")
boxplot(nst_data$NST.i.jaccard~groups_ag, xlab=" ", ylab="NST", ylim=c(0,1),yaxt='n')
axis(2, at=c(0,0.25,0.5,0.75,1))
box(which='plot')

nst_ba=nst_data$NST.i.jaccard[1:4]
nst_microbiome=nst_data$NST.i.jaccard[5:8]



mean(nst_ba)
sd(nst_ba)

mean(nst_microbiome)
sd(nst_microbiome)
```

```{r}

library(NST)

cg_filt2=subset_samples(cg_filt, definition !="bacterioplankton")

cg_filt2

try1=subset_samples(cg_filt2, Treatment3!="Bact_T0_filtered")
try1=subset_samples(try1, Treatment3!="Bact_T0")
try1=subset_samples(try1, treatment_26!="Bacteria Control T5")

strain=cbind( sample_data(try1)$diatom_strain)
strain=data.frame(strain)
row.names(strain)=sample_data(try1)$X.OTU.ID
colnames(strain)=c( "Strain") # metagroup - strain and group - BA_Micro

groupie=cbind( sample_data(try1)$definition)
groupie=data.frame(groupie)
row.names(groupie)=sample_data(try1)$X.OTU.ID
colnames(groupie)=c( "Treatment") # metagroup - strain and group - BA_Micro
dim(groupie)

otu_tab=t(otu_table(try1))
# taxonomic nst 
nst_1=tNST(comm = otu_tab , group = groupie, meta.group = strain, dist.method = 'jaccard',abundance.weighted=FALSE)
nst_data_ag=nst_1$index.grp
nst_data_ag
nst_data_group=nst_data_ag$ST.i.jaccard
names(nst_data_group)=c("co", "algal-cell")
par(mar=c(10,5,1,1))
groups_ag=c('co',"algal-cell")
barplot(nst_data_group,xlab=" ", ylab="NST", ylim=c(0,1),yaxt='n')
axis(2, at=c(0,0.25,0.5,0.75,1))
box(which='plot')

```

```{r}

library(NST)

cg_filt2=subset_samples(cg_filt, definition !="bacterioplankton")

cg_filt2

try1=subset_samples(cg_filt2, Treatment3!="Bact_T0_filtered")
try1=subset_samples(try1, Treatment3!="Bact_T0")
try1=subset_samples(try1, treatment_26!="Bacteria Control T5")

strain=cbind( sample_data(try1)$diatom_strain)
strain=data.frame(strain)
row.names(strain)=sample_data(try1)$X.OTU.ID
colnames(strain)=c( "Strain") # metagroup - strain and group - BA_Micro

groupie=cbind( sample_data(try1)$definition)
groupie=data.frame(groupie)
row.names(groupie)=sample_data(try1)$X.OTU.ID
colnames(groupie)=c( "Treatment") # metagroup - strain and group - BA_Micro
dim(groupie)

otu_tab=t(otu_table(try1))
# taxonomic nst 
nst_1=tNST(comm = otu_tab , group = groupie, meta.group = strain, dist.method = 'jaccard',abundance.weighted=FALSE)
nst_data_ag=nst_1$index.grp
nst_data_ag
nst_data_group=nst_data_ag$ST.i.jaccard
names(nst_data_group)=c("co", "algal-cell")
par(mar=c(7,5,1,1))
groups_ag=c('co',"algal-cell")
barplot(nst_data_group,xlab=" ", ylab="NST", ylim=c(0,1),yaxt='n', col=c("gray60","gray40"),las=2,cex.names=1.2)
axis(2, at=c(0,0.25,0.5,0.75,1))
#box(which='plot')

```
# Beta diversity and ordinations

## variance stabilizing transformation 

```{r vst-1, message=FALSE, error=FALSE, warning=FALSE}
library(CoDaSeq)
library(DESeq2)
cg_filt <- filter_taxa(cg_filt, function(x) sum(x) > 1, prune = TRUE)

########################################################################
# variance stabilizing transformation 
# from 16S stats 
ddst=phyloseq_to_deseq2(cg_filt,~Treatment3)
# 
gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
trial.deseq=estimateSizeFactors(ddst,geoMeans = apply(counts(ddst), 1, gm_mean))
data.vst <- DESeq2::varianceStabilizingTransformation(trial.deseq, blind = T)
# Apply variance stabilizing transformation
# Extract transformed OTU table
fieldEco.vstMat <- assay(data.vst)
# And put into my phloseq object
OTU = otu_table(fieldEco.vstMat, taxa_are_rows=T)
comp_vst=phyloseq(OTU,map,tax2,tree)
comp_clr=comp_vst
```

## all

### pca
```{r vst-2, message=FALSE, error=FALSE, warning=FALSE}
library(CoDaSeq)
library(DESeq2)
cg_filt2=subset_samples(cg_filt)
########################################################################
# variance stabilizing transformation 
# from 16S stats 
ddst=phyloseq_to_deseq2(cg_filt2,~Treatment3)
# 
gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
trial.deseq=estimateSizeFactors(ddst,geoMeans = apply(counts(ddst), 1, gm_mean))
data.vst <- DESeq2::varianceStabilizingTransformation(trial.deseq, blind = T)
# Apply variance stabilizing transformation
# Extract transformed OTU table
fieldEco.vstMat <- assay(data.vst)
# And put into my phloseq object
OTU = otu_table(fieldEco.vstMat, taxa_are_rows=T)
comp_vst=phyloseq(OTU,map,tax2,tree)
comp_clr=comp_vst
p=prcomp(t(otu_table(comp_clr)))

{plot(p$x[,1],p$x[,2],pch=sample_data(comp_clr)$pch,
     bg=as.character(sample_data(comp_clr)$col2),
     xlab = "PCA 42 %", ylab= "PCA 13 %")
ordiellipse(p$x,
     #    group=as.factor(sample_data(comp_clr)$Treatment),
          group=as.factor(sample_data(comp_clr)$treatment_26),
       #    col=c('gray',  'gray','#4D9DE0','#4D9DE0','gray',

     #     col=c('navy', 'firebrick', 'forestgreen','black'),
           kind ='sd',conf=0.8,
           label=TRUE)

}

```

### pcoa
```{r vst-2-pcoa, message=FALSE, error=FALSE, warning=FALSE}
otu2=otu_table(comp_clr)
euc=phyloseq::distance(t(otu2),'euclidean')
p=prcomp(euc,scale=T,center=T)

{
  
  par(mar=c(10,
            5,1,1),xpd=FALSE)
  plot(p$x[,1],p$x[,2],pch=sample_data(comp_clr)$pch,
     bg=as.character(sample_data(comp_clr)$col2),
     xlab = "PCoA1 65 %", 
     ylab= "PCoA2 21 %")
 # axis(2, at=c(-2,0,2))
ordiellipse(p$x,          group=as.factor(sample_data(comp_clr)$treatment_26),

           kind ='sd',conf=0.8,
           label=TRUE)
legend(-2, -5,
       legend=c("NbO-A5 BA","NbO-A5 Micro", "NbO-D4 BA", "NbO-D4 Micro", "NbP-YE5 BA", "NbP-YE5 Micro", "NbQ-B7 BA",
                "NbQ-B7 Micro"),
       bty='n',
       pt.bg=c("#4D9DE0", "#4D9DE0", "#E15554", "#E15554", "#3BB273", "#3BB273", "#7768AE", "#7768AE"),

       pch=c(21,22,21,22,21,22,21,22))

}


anosim(euc, grouping=as.factor(sample_data(comp_clr)$Pop3))
```

## microbiomes only 

### pca
```{r vst-3, message=FALSE, error=FALSE, warning=FALSE}
library(CoDaSeq)
library(DESeq2)
cg_filt2=subset_samples(cg_filt, extract=="WGA" & definition =='microbiome')
########################################################################
# variance stabilizing transformation 
# from 16S stats 
ddst=phyloseq_to_deseq2(cg_filt2,~Treatment3)
# 
gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
trial.deseq=estimateSizeFactors(ddst,geoMeans = apply(counts(ddst), 1, gm_mean))
data.vst <- DESeq2::varianceStabilizingTransformation(trial.deseq, blind = T)
# Apply variance stabilizing transformation
# Extract transformed OTU table
fieldEco.vstMat <- assay(data.vst)
# And put into my phloseq object
OTU = otu_table(fieldEco.vstMat, taxa_are_rows=T)
comp_vst=phyloseq(OTU,map,tax2,tree)
comp_clr=comp_vst
p=prcomp(t(otu_table(comp_clr)))
# summary(p)

{plot(p$x[,1],p$x[,2],pch=sample_data(comp_clr)$pch,
     bg=as.character(sample_data(comp_clr)$col),
     xlab = "PCA 24 %", ylab= "PCA 9 %")
ordiellipse(p$x,
          group=as.factor(sample_data(comp_clr)$treatment_26),
           kind ='sd',conf=0.8,
           label=TRUE)


}



```

### pcoa

```{r vst-4, message=FALSE, error=FALSE, warning=FALSE}
otu2=otu_table(comp_clr)
euc=phyloseq::distance(t(otu2),'euclidean')
p=prcomp(euc,scale=T,center=T)

# anosim(euc, grouping=as.factor(sample_data(comp_clr)$Pop3))
# anosim(euc, grouping=as.factor(sample_data(comp_clr)$diatom_control))

{
  
  par(mar=c(15,
            5,1,1),xpd=FALSE)
  plot(p$x[,1],p$x[,2],pch=sample_data(comp_clr)$pch,
     bg=as.character(sample_data(comp_clr)$col2),
     xlab = "PCoA1 36 %",
     ylab= "PCoA2 11 %", 
     yaxt='n', xaxt='n')
axis(2, at=c(-2,0,4))
axis(1, at=c(-4,0,5,10))
ordiellipse(p$x,
          group=as.factor(sample_data(comp_clr)$treatment_26),
           kind ='sd',conf=0.8,
           label=TRUE)

}

anosim(euc, grouping=as.factor(sample_data(comp_clr)$treatment_26))
adonis2(euc~as.factor(sample_data(comp_clr)$treatment_26))

# cg_filta=subset_samples(comp_clr, clonal=='monoclonal')
# otu2=otu_table(cg_filta)
# euc=phyloseq::distance(t(otu2),'euclidean')
# anosim(euc, grouping=as.factor(sample_data(cg_filta)$Pop3))
# anosim(euc, grouping=as.factor(sample_data(cg_filta)$Pop3))
# adonis2(euc~as.factor(sample_data(cg_filta)$Pop3))

```


## assemblages 

### pca
```{r vst-5, message=FALSE, error=FALSE, warning=FALSE}
library(CoDaSeq)
library(DESeq2)
cg_filt2=subset_samples(cg_filt, definition =='assemblage')
########################################################################
# variance stabilizing transformation 
# from 16S stats 
ddst=phyloseq_to_deseq2(cg_filt2,~Treatment3)
# 
gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
trial.deseq=estimateSizeFactors(ddst,geoMeans = apply(counts(ddst), 1, gm_mean))
data.vst <- DESeq2::varianceStabilizingTransformation(trial.deseq, blind = T)
# Apply variance stabilizing transformation
# Extract transformed OTU table
fieldEco.vstMat <- assay(data.vst)
# And put into my phloseq object
OTU = otu_table(fieldEco.vstMat, taxa_are_rows=T)
comp_vst=phyloseq(OTU,map,tax2,tree)
comp_clr=comp_vst


p=prcomp(t(otu_table(comp_clr)))
# summary(p)

{plot(p$x[,1],p$x[,2],pch=sample_data(comp_clr)$pch,
     bg=as.character(sample_data(comp_clr)$col),
     xlab = "PCA 28 %", ylab= "PCA 11 %")
ordiellipse(p$x,
          group=as.factor(sample_data(comp_clr)$treatment_26),
           kind ='sd',conf=0.8,
           label=TRUE)

}


```

### pcoa
```{r vst-6, message=FALSE, error=FALSE, warning=FALSE}
otu2=otu_table(comp_clr)
euc=phyloseq::distance(t(otu2),'euclidean')
p=prcomp(euc,scale=T,center=T)
# summary(p)

{
  
  par(mar=c(10,
            5,1,1),xpd=FALSE)
  plot(p$x[,1],p$x[,2],pch=sample_data(comp_clr)$pch,
     bg=as.character(sample_data(comp_clr)$col2),
     xlab = "PCoA1 41 %",
     ylab= "PCoA2 10 %",
     yaxt='n')
  axis(2, at=c(-2,0,2))
ordiellipse(p$x,
     #    group=as.factor(sample_data(comp_clr)$Treatment),
          group=as.factor(sample_data(comp_clr)$Pop3),
           col=c('black', 'navy', 'firebrick','forestgreen'),

     #     col=c('navy', 'firebrick', 'forestgreen','black'),
           kind ='sd',conf=0.8,
           label=TRUE)
legend(-2, -5,
       legend=c("RA5","RD4", "YC5", "NB7", "RD4 + RA5", "RD4 + YE5", "RD4+NB7"),
       bty='n',
  #    col=c("navy",'navy', 'firebrick', 'forestgreen', 'black','black','black'),
       pt.bg=c("navy",'navy', 'firebrick', 'forestgreen', 'black','black','black'),

       pch=c(22,23,22,23,21,22,23))

}


```

## assemblages with control 
```{r vst-7, message=FALSE, error=FALSE, warning=FALSE}
library(CoDaSeq)
library(DESeq2)
cg_filt2=subset_samples(cg_filt, definition =='assemblage' | definition == "inoculum_T5")
########################################################################
# variance stabilizing transformation 
# from 16S stats 
ddst=phyloseq_to_deseq2(cg_filt2,~Treatment3)
# 
gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
trial.deseq=estimateSizeFactors(ddst,geoMeans = apply(counts(ddst), 1, gm_mean))
data.vst <- DESeq2::varianceStabilizingTransformation(trial.deseq, blind = T)
# Apply variance stabilizing transformation
# Extract transformed OTU table
fieldEco.vstMat <- assay(data.vst)
# And put into my phloseq object
OTU = otu_table(fieldEco.vstMat, taxa_are_rows=T)
comp_vst=phyloseq(OTU,map,tax2,tree)
comp_clr=comp_vst

otu2=otu_table(comp_clr)
euc=phyloseq::distance(t(otu2),'euclidean')
p=prcomp(euc,scale=T,center=T)


{
  
  par(mar=c(15,
            5,1,1),xpd=FALSE)
  plot(p$x[,1],p$x[,2],pch=sample_data(comp_clr)$pch,
     bg=as.character(sample_data(comp_clr)$col2),
     xlab = "PCoA1 41.11 %",
     ylab= "PCoA2 10.18 %",
     yaxt='n')
  axis(2, at=c(-2,0,2))
ordiellipse(p$x,
          group=as.factor(sample_data(comp_clr)$treatment_26),
           kind ='sd',conf=0.8,
           label=TRUE)


}



```


## assemblages with microbiome 
```{r vst-8, message=FALSE, error=FALSE, warning=FALSE}
library(CoDaSeq)
library(DESeq2)
cg_filt2=subset_samples(cg_filt, definition =='assemblage' | definition == "microbiome" | definition == "inoculum_T5")
########################################################################
# variance stabilizing transformation 
# from 16S stats 
ddst=phyloseq_to_deseq2(cg_filt2,~Treatment3)
# 
gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
trial.deseq=estimateSizeFactors(ddst,geoMeans = apply(counts(ddst), 1, gm_mean))
data.vst <- DESeq2::varianceStabilizingTransformation(trial.deseq, blind = T)
# Apply variance stabilizing transformation
# Extract transformed OTU table
fieldEco.vstMat <- assay(data.vst)
# And put into my phloseq object
OTU = otu_table(fieldEco.vstMat, taxa_are_rows=T)
comp_vst=phyloseq(OTU,map,tax2,tree)
comp_clr=comp_vst

otu2=otu_table(comp_clr)
euc=phyloseq::distance(t(otu2),'euclidean')
p=prcomp(euc,scale=T,center=T)


{
  
  par(mar=c(15,
            5,1,1),xpd=FALSE)
  plot(p$x[,1],p$x[,2],pch=sample_data(comp_clr)$pch,
     bg=as.character(sample_data(comp_clr)$col2),
     xlab = "PCoA1 75 %",
     ylab= "PCoA2 6 %",cex=1.5,yaxt='n')
  axis(2, at=c(-2,0,6))
ordiellipse(p$x,
          group=as.factor(sample_data(comp_clr)$treatment_26),
           kind ='sd',conf=0.8,
           label=FALSE)
legend('topright',
       legend=c("NbO-A5 BA","NbO-A5 Micro", "NbO-D4 BA", "NbO-D4 Micro", "NbP-YE5 BA", "NbP-YE5 Micro", "NbQ-B7 BA",
                "NbQ-B7 Micro", "Bacterial Control"),
       bty='n',
       pt.bg=c("#4D9DE0", "#4D9DE0", "#E15554", "#E15554", "#3BB273", "#3BB273", "#7768AE", "#7768AE",'gray'),

       pch=c(21,22,21,22,21,22,21,22,21))


}



```


## assemblages with microbiomeno contorl
```{r vst-9, message=FALSE, error=FALSE, warning=FALSE}
library(CoDaSeq)
library(DESeq2)
cg_filt2=subset_samples(cg_filt, definition =='assemblage' | definition == "microbiome" )
########################################################################
# variance stabilizing transformation 
# from 16S stats 
ddst=phyloseq_to_deseq2(cg_filt2,~Treatment3)
# 
gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
trial.deseq=estimateSizeFactors(ddst,geoMeans = apply(counts(ddst), 1, gm_mean))
data.vst <- DESeq2::varianceStabilizingTransformation(trial.deseq, blind = T)
# Apply variance stabilizing transformation
# Extract transformed OTU table
fieldEco.vstMat <- assay(data.vst)
# And put into my phloseq object
OTU = otu_table(fieldEco.vstMat, taxa_are_rows=T)
comp_vst=phyloseq(OTU,map,tax2,tree)
comp_clr=comp_vst

otu2=otu_table(comp_clr)
euc=phyloseq::distance(t(otu2),'euclidean')
p=prcomp(euc,scale=T,center=T)


{
  
  par(mar=c(15,
            5,1,1),xpd=FALSE)
  plot(p$x[,1],p$x[,2],pch=sample_data(comp_clr)$pch,
     bg=as.character(sample_data(comp_clr)$col2),
     xlab = "PCoA1 75 %",
     ylab= "PCoA2 6 %",cex=1.5,yaxt='n')
  axis(2, at=c(-2,0,6))
ordiellipse(p$x,
          group=as.factor(sample_data(comp_clr)$treatment_26),
           kind ='sd',conf=0.8,
           label=FALSE)
legend('bottomleft', ncol=2,
       legend=c("NbO-A5 BA","NbO-A5 Micro", "NbO-D4 BA", "NbO-D4 Micro", "NbP-YE5 BA", "NbP-YE5 Micro", "NbQ-B7 BA",
                "NbQ-B7 Micro"),
       bty='n',
       pt.bg=c("#4D9DE0", "#4D9DE0", "#E15554", "#E15554", "#3BB273", "#3BB273", "#7768AE", "#7768AE"),

       pch=c(21,22,21,22,21,22,21,22,21))


}
anosim(euc, grouping=sample_data(comp_clr)$definition)
anosim(euc, grouping=sample_data(comp_clr)$diatom_strain)

adonis2(euc ~ sample_data(comp_clr)$definition)

```

# philr 

### all

```{r philr-1, message=FALSE, error=FALSE, warning=FALSE}
library(philr)
sub=subset_samples(cg_filt, extract=="WGA" & definition =="microbiome")

GP <- transform_sample_counts(sub, function(x) x+1)
phy_tree(GP) <- makeNodeLabel(phy_tree(GP), method="number", prefix='n')
name.balance(phy_tree(GP), tax_table(GP), 'n1')

otu.table <- t(otu_table(GP))
tree <- phy_tree(GP)
metadata <- sample_data(GP)
tax <- tax_table(GP)

gp.philr <- philr(otu.table, tree, 
                  part.weights='enorm.x.gm.counts', 
                  ilr.weights='blw.sqrt')

gp.dist <- dist(gp.philr, method="euclidean")
gp.pcoa <- ordinate(GP, 'PCoA', distance=gp.dist)
gp.p=prcomp(gp.dist)

{
  par(mar=c(15,
            5,1,1),xpd=F)
  plot(gp.p$x[,1],gp.p$x[,2],pch=sample_data(sub)$pch,
     bg=as.character(sample_data(sub)$col2),
       xlab = "PCoA1 38.52 %",
     ylab= "PCoA2 19.90 %")
# axis(2, at=c(-2,0,2))
ordiellipse(gp.p$x,
    #    group=as.factor(sample_data(comp_clr)$Treatment),
         group=as.factor(sample_data(sub)$Pop3),
           col=c('black', 'navy', 'firebrick','forestgreen'),

     #     col=c('navy', 'firebrick', 'forestgreen','black'),
           kind ='sd',conf=0.8,
           label=TRUE)
legend(-20,-40,
        legend=c("RA5","RD4", "YC5", "NB7", "RD4 + RA5", "RD4 + YE5", "RD4 + NB7"),
       bty='n',
  #    col=c("navy",'navy', 'firebrick', 'forestgreen', 'black','black','black'),
       pt.bg=c("navy",'navy', 'firebrick', 'forestgreen', 'black','black','black'),

       pch=c(22,23,22,23,21,22,23))

}

```

### microbiomes monoclonal
```{r philr-2, message=FALSE, error=FALSE, warning=FALSE}
library(philr)
sub=subset_samples(cg_filt, extract=="WGA" & definition =="microbiome")

GP <- transform_sample_counts(sub, function(x) x+1)
phy_tree(GP) <- makeNodeLabel(phy_tree(GP), method="number", prefix='n')
name.balance(phy_tree(GP), tax_table(GP), 'n1')

otu.table <- t(otu_table(GP))
tree <- phy_tree(GP)
metadata <- sample_data(GP)
tax <- tax_table(GP)

gp.philr <- philr(otu.table, tree, 
                  part.weights='enorm.x.gm.counts', 
                  ilr.weights='blw.sqrt')

gp.dist <- dist(gp.philr, method="euclidean")
gp.pcoa <- ordinate(GP, 'PCoA', distance=gp.dist)
gp.p=prcomp(gp.dist)

{
  par(mar=c(4,
            5,1,1),xpd=F)
  plot(gp.p$x[,1],gp.p$x[,2],pch=sample_data(sub)$pch,
     bg=as.character(sample_data(sub)$col2),
       xlab = "PCoA1 39.68 %",
     ylab= "PCoA2 19.46 %")
# axis(2, at=c(-2,0,2))
ordiellipse(gp.p$x,
    #    group=as.factor(sample_data(comp_clr)$Treatment),
         group=as.factor(sample_data(sub)$treatment_26),
           col=c( 'navy', 'firebrick','forestgreen'),

     #     col=c('navy', 'firebrick', 'forestgreen','black'),
           kind ='sd',conf=0.8,
           label=TRUE)
legend(-20,-40,
        legend=c("RA5","RD4", "YC5", "NB7", "RD4 + RA5", "RD4 + YE5", "RD4 + NB7"),
       bty='n',
  #    col=c("navy",'navy', 'firebrick', 'forestgreen', 'black','black','black'),
       pt.bg=c("navy",'navy', 'firebrick', 'forestgreen'),

       pch=c(22,23,22,23,21,22,23))

}
set.seed(1234)
adonis2(gp.dist~(sample_data(sub)$treatment_26))
anosim(gp.dist,(sample_data(sub)$treatment_26))

```


### assemblage control 
```{r philr-3, message=FALSE, error=FALSE, warning=FALSE}
library(philr)
sub=subset_samples(cg_filt, definition =='assemblage' | definition == "inoculum_T5")
GP <- transform_sample_counts(sub, function(x) x+1)
phy_tree(GP) <- makeNodeLabel(phy_tree(GP), method="number", prefix='n')
name.balance(phy_tree(GP), tax_table(GP), 'n1')

otu.table <- t(otu_table(GP))
tree <- phy_tree(GP)
metadata <- sample_data(GP)
tax <- tax_table(GP)

gp.philr <- philr(otu.table, tree, 
                  part.weights='enorm.x.gm.counts', 
                  ilr.weights='blw.sqrt')

gp.dist <- dist(gp.philr, method="euclidean")
gp.pcoa <- ordinate(GP, 'PCoA', distance=gp.dist)
gp.p=prcomp(gp.dist)

{plot(gp.pcoa$vectors[,1],gp.pcoa$vectors[,2],pch=sample_data(sub)$pch,
     bg=as.character(sample_data(sub)$col),
     xlab = "PCoA 1 53 %", ylab= "PCoA 2 28%")
ordiellipse(gp.pcoa$vectors,
           group=as.factor(sample_data(sub)$Treatment),label=T,
           col=c('gray70','gray50','forestgreen','green','navy','blue'),
           kind ='sd',conf=0.8)}

```

### good
```{r philr-4, message=FALSE, error=FALSE, warning=FALSE}



library(philr)
sub=subset_samples(cg_filt, definition =='assemblage' | definition == "microbiome" )
GP <- transform_sample_counts(sub, function(x) x+1)
phy_tree(GP) <- makeNodeLabel(phy_tree(GP), method="number", prefix='n')
name.balance(phy_tree(GP), tax_table(GP), 'n1')

otu.table <- t(otu_table(GP))
tree <- phy_tree(GP)
metadata <- sample_data(GP)
tax <- tax_table(GP)

gp.philr <- philr(otu.table, tree, 
                  part.weights='enorm.x.gm.counts', 
                  ilr.weights='blw.sqrt')

gp.dist <- dist(gp.philr, method="euclidean")
gp.pcoa <- ordinate(GP, 'PCoA', distance=gp.dist)
gp.p=prcomp(gp.dist)

{
  par(mar=c(15,
            5,1,1),xpd=F)
  plot(gp.p$x[,1],gp.p$x[,2],pch=sample_data(sub)$pch,
     bg=as.character(sample_data(sub)$col2),
       xlab = "PCoA1 41  %",
     ylab= "PCoA2 26 %")
# axis(2, at=c(-2,0,2))
ordiellipse(gp.p$x,
    #    group=as.factor(sample_data(comp_clr)$Treatment),
         group=as.factor(sample_data(sub)$definition),

     #     col=c('navy', 'firebrick', 'forestgreen','black'),
           kind ='sd',conf=0.8,
           label=TRUE)
}



```

#### used in 
```{r philr-5, message=FALSE, error=FALSE, warning=FALSE}
library(CoDaSeq)
library(DESeq2)
cg_filt2=subset_samples(cg_filt, definition =='assemblage' | definition == "microbiome" )
########################################################################
# variance stabilizing transformation 
# from 16S stats 
ddst=phyloseq_to_deseq2(cg_filt2,~Treatment3)
# 
gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
trial.deseq=estimateSizeFactors(ddst,geoMeans = apply(counts(ddst), 1, gm_mean))
data.vst <- DESeq2::varianceStabilizingTransformation(trial.deseq, blind = T)
# Apply variance stabilizing transformation
# Extract transformed OTU table
fieldEco.vstMat <- assay(data.vst)
# And put into my phloseq object
OTU = otu_table(fieldEco.vstMat, taxa_are_rows=T)
comp_vst=phyloseq(OTU,map,tax2,tree)
comp_clr=comp_vst

otu2=otu_table(comp_clr)
euc=phyloseq::distance(t(otu2),'euclidean')
p=prcomp(euc,scale=T,center=T)


library(philr)
sub=subset_samples(cg_filt, definition =='assemblage' | definition == "microbiome" )
GP <- transform_sample_counts(sub, function(x) x+1)
phy_tree(GP) <- makeNodeLabel(phy_tree(GP), method="number", prefix='n')
name.balance(phy_tree(GP), tax_table(GP), 'n1')

otu.table <- t(otu_table(GP))
tree <- phy_tree(GP)
metadata <- sample_data(GP)
tax <- tax_table(GP)

gp.philr <- philr(otu.table, tree, 
                  part.weights='enorm.x.gm.counts', 
                  ilr.weights='blw.sqrt')

gp.dist <- dist(gp.philr, method="euclidean")
gp.pcoa <- ordinate(GP, 'PCoA', distance=gp.dist)
gp.p=prcomp(gp.dist)


{
  par(mar=c(5,
            5,1,1),xpd=F,mfrow=c(1,2))

  plot(p$x[,1],p$x[,2],pch=sample_data(comp_clr)$pch,
       bg=as.character(sample_data(comp_clr)$col2),
       xlab = "PCoA1 75 %",
       ylab= "PCoA2 6 %",cex=1.5,yaxt='n', xaxt='n')
  axis(1, at=c(-8,-4,0,4))
  axis(2, at=c(-2,0,6))
  ordiellipse(p$x,
              group=as.factor(sample_data(comp_clr)$treatment_26),
              kind ='sd',conf=0.8,
              label=FALSE)
 # legend('bottomleft', ncol=2,
  #       legend=c("NbO-A5 BA","NbO-A5 Micro", "NbO-D4 BA", "NbO-D4 Micro", "NbP-YE5 BA", "NbP-YE5 Micro", "NbQ-B7 BA",
   #               "NbQ-B7 Micro"),
 #        bty='n',
  #       pt.bg=c("#4D9DE0", "#4D9DE0", "#E15554", "#E15554", "#3BB273", "#3BB273", "#7768AE", "#7768AE"),
  #       
   #      pch=c(21,22,21,22,21,22,21,22,21))
  
  plot(gp.p$x[,1],gp.p$x[,2],pch=sample_data(sub)$pch,
       bg=as.character(sample_data(sub)$col2),
       xlab = "PCoA1 41  %",cex=1.5,
       xaxt='n', yaxt='n',
       ylab= "PCoA2 26 %")
axis(1, at=c(-20,0,40))
axis(2, at=c(-20,0,20))
ordiellipse(gp.p$x,
              #    group=as.factor(sample_data(comp_clr)$Treatment),
              group=as.factor(sample_data(sub)$treatment_26),
              
              #     col=c('navy', 'firebrick', 'forestgreen','black'),
              kind ='sd',conf=0.8,
              label=FALSE)
}


anosim(gp.dist, grouping=sample_data(sub)$definition)
anosim(gp.dist, grouping=sample_data(sub)$diatom_strain)

adonis2(gp.dist ~ sample_data(sub)$definition)
adonis2(euc ~ sample_data(sub)$definition)

adonis2(gp.dist ~ sample_data(sub)$diatom_strain + sample_data(sub)$definition, by='margin')

adonis2(euc ~ sample_data(sub)$diatom_strain + sample_data(sub)$definition, by='margin')


```

# pref index

```{r}
#agg1=aggregate_taxa(cg_filt, "species")
#write.csv(otu_table(agg1), "agg_genus_otu.csv")


```
